---
titlK: "Germination modelling"
author: "Marion Carme"
datK: "2023-04-13"
output:
  html_document:
    df_print: paged
  latex_enginK: xelatex
  pdf_document: 
editor_options:
  chunk_output_typK: console
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r packages}
library(anyLib)#to load packages and install them if not already installed.
anyLib(list("readxl",
            "here",
            "tidyverse",
            "survival",#to use Cox Proportional Hazards Regression Models
            "survminer",#Contains the function 'ggsurvplot()' for drawing easily beautiful and 'ready-to-publish' survival curves with the 'number at risk' table and 'censoring count plot'.
            "coxme",#Mixed cox models
            "gridExtra",#pour utiliser grid.arrange
            "ggtext",#pour utiliser element_markdown
            "performance",#Pour évaluer les modèles
            "lme4",#Pour les modèles linéaire généralisés mixtes
            "FactoMineR",#Statistiques exploratoires
            "factoextra",#Statistiques exploratoires
            "SurvMetrics",#La fonction CoxCompare calcule différents indices avec Survmetrics (most of the evaluation metrics for survival predictions)
            "pec",
            "ciTools",#Confidence or Prediction Intervals, Quantiles, and Probabilities for Statistical
            "autoReg",#to extract results from survival fit
            "emmeans",#Quantification of differences between groups in a model
            "car",#statistiques
            "MuMIn",#Evaluation and comparing models
            "DHARMa",#residual diagnostics for hierarchical (multi-level/mixed) regression models #citation("DHARMa")
            "glmmTMB",#To make zero inflated negative binomial regression
            "raster",
            "rworldxtra",
            "rworldmap",
            "rnaturalearth",
            "rnaturalearthdata",
            "stringr",
            "sp",
            "sf",
            "germinationmetrics",#Calculations of germination indexes
            "corrr",#function correlate
            "ggnetwork",#to do traits correlation networks
            "network",#to do traits correlation networks
            "Hmisc",#rcorr for p value
            "betareg",#for beta regression
            "RVAideMemoire", #some functions coming from the guide of Hervé Maxime
            "multcomp",#multiple comparison
            "rsq",#R² of poisson GLM
            "xlsx",#to export xlsx with several sheets
            "metrica",#for scatterplot cross validation
            "tidytext",#to use reorder_within because adding an interaction to betareg regression was making the graph unordered
            "rstatix",#doing post hoc comparison on curves
            "rcompanion",#use cldList for having letters when doing post hoc comparison on curves
            "ggrepel",#utiliser geom_textrepel
            "interactions",#utiliser interact_plot pour afficher les interactions d'un modèle
            "condformat",#Format tables
            "ggh4x",#to use facet_wrap2 and color facets
            "DMwR",#unscale
            "ggspatial",#north arrow on the map
            "rgl",#plot3D
            "jtools"#interact_plot
            ))
#remotes::install_version("DMwR", version="0.4.1")
```

#Climate Data import

We extracted climatic data for each population, relying on their location, using the [Climate Downscaling Tool (ClimateDT)](https://www.ibbr.cnr.it/climate-dt/) developed within the framework of the B4EST project. Climatic variables have been extracted from 1901 to 2022.

```{r Data import climate}
#Can give up on mother since all its variables are correlated with genetic effect variables
ClimateGenetBDD <- (read_excel(here("Climate_Provenances.xlsx"), sheet = "Genetic") %>% 
  rename("Provenance"="ID"))[,c("Provenance","Tmax_spring","Pmean_spring","Tmean_annual","Trange_day","Isothermality","Tseasonality","Trange_annual","Pmean_annual","Pmax_annual", "Pmin_annual", "Pseasonality","Evapotranspiration", "ClimaticMoistureDeficit","Tmin_annual","Pmean_summer","Tmax_annual", "ChillingDD0", "GrowingDD5", "CoolingDD18", "FrostFreeDays", "DM_AridityIndex","E_PluviometricQuotient","RM_ThermicityIndex","AHM","SHM","TD","Tmin_30years")] %>%
  group_by(Provenance) %>% 
  summarise_all(mean)
ClimateAcmatBDD <- (read_excel(here("Climate_Provenances.xlsx"),sheet = "AcornMaturation") %>% 
  rename("Provenance"="ID"))[,c("Provenance","Tmax_spring","Pmean_spring","Tmean_annual","Trange_day","Isothermality","Tseasonality","Trange_annual","Pmean_annual","Pmax_annual", "Pmin_annual", "Pseasonality","Evapotranspiration", "ClimaticMoistureDeficit","Tmin_annual","Pmean_summer","Tmax_annual", "ChillingDD0", "GrowingDD5", "CoolingDD18", "FrostFreeDays", "DM_AridityIndex","E_PluviometricQuotient","RM_ThermicityIndex","AHM","SHM","TD","Tmin_30years")] %>% 
  group_by(Provenance) %>% 
  summarise_all(mean)
colnames(ClimateAcmatBDD) <- paste0(colnames(ClimateAcmatBDD),"_Mat")
ClimateMTreeBDD <- (read_excel(here("Climate_Provenances.xlsx"),sheet = "Mother") %>% 
  rename("Provenance"="ID"))[,c("Provenance","Tmax_spring","Pmean_spring","Tmean_annual","Trange_day","Isothermality","Tseasonality","Trange_annual","Pmean_annual","Pmax_annual", "Pmin_annual", "Pseasonality","Evapotranspiration", "ClimaticMoistureDeficit","Tmin_annual","Pmean_summer","Tmax_annual", "ChillingDD0", "GrowingDD5", "CoolingDD18", "FrostFreeDays", "DM_AridityIndex","E_PluviometricQuotient","RM_ThermicityIndex","AHM","SHM","TD","Tmin_30years")] %>% 
  group_by(Provenance) %>% 
  summarise_all(mean)
colnames(ClimateMTreeBDD) <- paste0(colnames(ClimateMTreeBDD),"_MTree")
ClimateProv = ClimateGenetBDD %>% 
  left_join(ClimateAcmatBDD %>% rename("Provenance" = "Provenance_Mat"),by = "Provenance") %>% 
  left_join(ClimateMTreeBDD %>% rename("Provenance" = "Provenance_MTree"),by = "Provenance") %>% 
  mutate(Population = case_when(Provenance == "Moliets et Maa" ~ "FRA_MoeMa",
                                Provenance == "Padouen" ~ "FRA_Pad",
                                Provenance == "Fuencaliente" ~ "SPA_Fue",
                                Provenance == "Tuscania" ~ "ITA_Tus",
                                Provenance == "Quinta de Serra" ~ "POR_QudSe",
                                Provenance == "Tabarka" ~ "TUN_Tab",
                                Provenance == "Fernana" ~ "TUN_Fer",
                                Provenance == "Los Alcornocales" ~ "SPA_LoAl")) 
ClimateProvScaled = ClimateProv %>% 
  mutate_at(c("Tmax_spring","Pmean_spring","Tmean_annual","Trange_day","Isothermality","Tseasonality","Trange_annual","Pmean_annual","Pmax_annual", "Pmin_annual", "Pseasonality","Evapotranspiration", "ClimaticMoistureDeficit","Tmin_annual","Pmean_summer","Tmax_annual", "ChillingDD0", "GrowingDD5", "CoolingDD18", "FrostFreeDays", "DM_AridityIndex","E_PluviometricQuotient","RM_ThermicityIndex","AHM","SHM","TD","Tmin_30years",
              "Tmax_spring_Mat", "Pmean_spring_Mat", "Tmean_annual_Mat", "Trange_day_Mat", "Isothermality_Mat", "Tseasonality_Mat", "Trange_annual_Mat", "Pmean_annual_Mat", "Pmax_annual_Mat", "Pmin_annual_Mat", "Pseasonality_Mat", "Evapotranspiration_Mat","ClimaticMoistureDeficit_Mat", "Tmin_annual_Mat", "Pmean_summer_Mat", "Tmax_annual_Mat", "ChillingDD0_Mat", "GrowingDD5_Mat", "CoolingDD18_Mat", "FrostFreeDays_Mat", "DM_AridityIndex_Mat", "E_PluviometricQuotient_Mat", "RM_ThermicityIndex_Mat", "AHM_Mat", "SHM_Mat","TD_Mat","Tmin_30years_Mat",
              "Tmax_spring_MTree", "Pmean_spring_MTree", "Tmean_annual_MTree", "Trange_day_MTree", "Isothermality_MTree", "Tseasonality_MTree", "Trange_annual_MTree", "Pmean_annual_MTree", "Pmax_annual_MTree", "Pmin_annual_MTree", "Pseasonality_MTree", "Evapotranspiration_MTree","ClimaticMoistureDeficit_MTree", "Tmin_annual_MTree", "Pmean_summer_MTree", "Tmax_annual_MTree", "ChillingDD0_MTree", "GrowingDD5_MTree", "CoolingDD18_MTree", "FrostFreeDays_MTree", "DM_AridityIndex_MTree", "E_PluviometricQuotient_MTree", "RM_ThermicityIndex_MTree", "AHM_MTree", "SHM_MTree","TD_MTree","Tmin_30years_MTree"), funs(c(scale(.))))
```

Rivas-Martinex thermicity index: This is the sum in tenths of a degree de T (average annual temperature), m (average temperature of the minimums of the coldest month) and M (average temperature of the maximums of the coldest monthly period) It = (T+m+M) 10. It is therefore an index that considers the intensity of the cold, a limiting factor for many plants and plant communities, together with the average annual temperature. The correlation between the values of this index and the vegetation is quite satisfactory in warm and temperate climates. In cold climates with It or Itc values of less than 120, and in contental climates with an Ic > 21, it is more significant and precise to use the value of the positive annual temperature (Tp).

Hargreaves climatic moisture deficit: sum of the monthly difference between a reference evaporation (Eref) and precipitation.

De Martonne aridity index (1926): reflects the aridity of the climate, which combines average temperature and precipitation (the influence of temperature appears to predominate at high latitudes, while that of rainfall predominates near the equator). It allows us to determine what constitutes a dry climate, a dry month, etc., something that cannot be done using precipitation variables alone.

Emberger’s Precipitation Quotient: Louis Emberger also proposed calculating a quotient, an empirical measure of rainfall effectiveness. This precipitation quotient, or Emberger climate index, is used to define the five different types of Mediterranean climates, ranging from the driest to the high-mountain climate. Emberger’s index (1932) defines the degree of humidity in a climate. It takes into account annual precipitation P, the average of the maximum temperatures for the hottest month (M), and the average of the minimum temperatures for the coldest month (m). As with Gaussen’s xerothermic index, it is particularly well-suited to Mediterranean regions, where it allows for the distinction of different climatic zones. In these regions, Emberger observed that the temperature range (M–m)—and thus evaporation—is an important factor in plant distribution. It is well known that, at the same average temperature, evaporation increases as the temperature range increases. The precipitation factor taken into account is the product of the number of rainy days per year (n) and the average annual precipitation (P).

Annual heat moisture index: measure of aridity, Tmean_annual, Pmean_annual
Summer heat moisture index: aridity in summer, Tmax_annual, Pmean_summer

#Reducing climatic variable number

We chose to select primarily the climatic indexes that are more integrated and take into account several variables.

ClimaticMoistureDeficit(evap, Pmean_annual)
RM_ThermicityIndex (Tmean_annual, Tmin_annual) (but SHM and Pseasonality not the same even if correlated? - that's ok if there are not significant)
DM_AridityIndex ou Pmean_spring?
E_PluviometricQuotient
Tseasonality
Tmax_spring

```{r  ACP correlation climatic variables mother VS Acorn maturation VS provenance}
#ACP correlation climatic variables mother VS provenance
ACP = PCA(ClimateProvScaled[,c("Provenance","ClimaticMoistureDeficit","Tmax_spring", "RM_ThermicityIndex", "DM_AridityIndex","E_PluviometricQuotient","Tseasonality", "SHM","Pseasonality","ClimaticMoistureDeficit_MTree","Tmax_spring_MTree", "RM_ThermicityIndex_MTree", "DM_AridityIndex_MTree","E_PluviometricQuotient_MTree","Tseasonality_MTree", "SHM_MTree","Pseasonality_MTree")], 
          scale.unit =TRUE, ncp = 6, graph = F, quali.sup = c(1))#, quanti.sup=c(6,10), quali.sup = c(1,3)
grid.arrange(
  plot.PCA(ACP,axes=c(1,2), choix = "ind", habillage = "none", col.ind = "black",col.ind.sup="blue",col.quali = "magenta",label = c("ind","ind.sup","quali"), select = "cos2 10",unselect=0.9),#graphe individus
  fviz_pca_var(ACP, axes = c(1, 2), col.quanti.sup = "blue", col.circle = "grey70",repel=T),#graphe variables
  ncol = 2)
fviz_eig(ACP, addlabels = TRUE, ylim = c(0, 50))#Graphe des éboulis
cor(ClimateProvScaled[,c("ClimaticMoistureDeficit","Tmax_spring", "RM_ThermicityIndex", "DM_AridityIndex","E_PluviometricQuotient","Tseasonality", "SHM","Pseasonality","ClimaticMoistureDeficit_MTree","Tmax_spring_MTree", "RM_ThermicityIndex_MTree", "DM_AridityIndex_MTree","E_PluviometricQuotient_MTree","Tseasonality_MTree", "SHM_MTree","Pseasonality_MTree")])

#ACP correlation climatic variables acorn maturation VS provenance
ACP = PCA(ClimateProvScaled[,c("Provenance","ClimaticMoistureDeficit","Tmax_spring", "RM_ThermicityIndex", "DM_AridityIndex","E_PluviometricQuotient","Tseasonality", "SHM","Pseasonality","ClimaticMoistureDeficit_Mat","Tmax_spring_Mat", "RM_ThermicityIndex_Mat", "DM_AridityIndex_Mat","E_PluviometricQuotient_Mat","Tseasonality_Mat", "SHM_Mat","Pseasonality_Mat")], 
          scale.unit =TRUE, ncp = 6, graph = F, quali.sup = c(1))#, quanti.sup=c(6,10), quali.sup = c(1,3)
grid.arrange(
  plot.PCA(ACP,axes=c(1,2), choix = "ind", habillage = "none", col.ind = "black",col.ind.sup="blue",col.quali = "magenta",label = c("ind","ind.sup","quali"), select = "cos2 10",unselect=0.9),#graphe individus
  fviz_pca_var(ACP, axes = c(1, 2), col.quanti.sup = "blue", col.circle = "grey70",repel=T),#graphe variables
  ncol = 2)
fviz_eig(ACP, addlabels = TRUE, ylim = c(0, 50))#Graphe des éboulis
cor(ClimateProvScaled[,c("ClimaticMoistureDeficit","Tmax_spring", "RM_ThermicityIndex", "DM_AridityIndex","E_PluviometricQuotient","Tseasonality", "SHM","Pseasonality","ClimaticMoistureDeficit_Mat","Tmax_spring_Mat", "RM_ThermicityIndex_Mat", "DM_AridityIndex_Mat","E_PluviometricQuotient_Mat","Tseasonality_Mat", "SHM_Mat","Pseasonality_Mat")],)
```
Between Provenance variables and n and n-1 variables:
All correlated except: Pmean_spring, Pmax_annual, and a little bit Pseasonality
Between Provenance variables and mother tree variables:
All correlated

We chose to work with the 1901-1960 period, i.e. the long-term climate of the population, to which the population may have adapted to. 

```{r  ACP choose non correlated variable}
ACP = PCA(ClimateProvScaled[,c("Provenance","ClimaticMoistureDeficit", "Tmax_spring", "RM_ThermicityIndex", "DM_AridityIndex", "E_PluviometricQuotient", "Tseasonality")], scale.unit =TRUE, ncp = 6, graph = F, quali.sup = c(1))#, quanti.sup=c(6,10), quali.sup = c(1,3)
grid.arrange(
  plot.PCA(ACP,axes=c(1,2), choix = "ind", habillage = "none", col.ind = "black",col.ind.sup="blue",col.quali = "magenta",label = c("ind","ind.sup","quali"), select = "cos2 10",unselect=0.9),#graphe individus
  fviz_pca_var(ACP, axes = c(1, 2), col.quanti.sup = "blue", col.circle = "grey70"),#graphe variables
  #fviz_pca_var(ACP, axes = c(2, 3), col.quanti.sup = "blue", col.circle = "grey70"),#graphe variables
  #fviz_pca_var(ACP, axes = c(3, 4), col.quanti.sup = "blue", col.circle = "grey70"),#graphe variables
  ncol = 2)
fviz_eig(ACP, addlabels = TRUE, ylim = c(0, 50))#Graphe des éboulis
```

In order to select the most explanatory index between Hargreaves' climatic water deficit (Hargreaves & Samani, 1985) and De Martonne's aridity index (De Martonne, 1926) - which were highly negatively correlated - we fitted two models for each of the three germination traits, each model incorporating the selected climatic variables and either Hargreaves' climatic water deficit (Hargreaves & Samani, 1985) or De Martonne's aridity index (De Martonne, 1926), and carried out model comparisons using Akaike's information criterion (AIC). 

1ère sélection: Tseasonality + Pmean_summer + Pmean_spring + Pmax_annual + Tmin_annual + Pseasonality + Tmax_spring + Pmean_spring_Mat + Pmax_annual_Mat
2ème sélection: "ClimaticMoistureDeficit", "Tmax_spring", "RM_ThermicityIndex", "DM_AridityIndex", "Tseasonality", "SHM", "Pseasonality"

Then, still multicollinearity, choices to make:
Aridity Index or Pluviometric Quotient
ClimaticMoistureDeficit or Tseasonality
Aridity Index or Climatic Moisture Deficit

We remove Pluviometric Quotient, because increase vif in all models.
Aridity index and climatic moisture deficit can be ok in the same model, but there are anticorrelated so they still bring the same kind of information.
FINAL CHOICK:
Water stress + spring temperature + temperature seasonality + Cold intensity
ClimaticMoistureDeficit + Tmax_spring + Tseasonality + RM_ThermicityIndex
We choose CMD over aridity index because CMD explain more germination than aridity index (models compared).


#Color code

The color code (which color for which population) has been done according to the population variable that was shaping the best germination dynamic. Cox models have been built, then run to do predictions. Several color codes have tested for each population variable, and summaries and Anova table of the models have been analyzed.
This led to the following color code, according to Rivaz-Martinez Thermicity index.

```{r Which colors to summarise/represent populations}
RM_ThermicityIndex_tab = data.frame(unique(ClimateProv[,c("Population","RM_ThermicityIndex")])) %>% 
  arrange(-RM_ThermicityIndex)
#Climate_tab = data.frame(unique(ClimateProv[,c("Population","DM_AridityIndex")])) %>% 
#arrange(-DM_AridityIndex) %>% 
#mutate(DM_AridityQuali = ifelse(DM_AridityIndex>20 & DM_AridityIndex<30,"Subhumid",ifelse(DM_AridityIndex>30 & DM_AridityIndex<60,"Humid","")))
#CMD_tab = data.frame(unique(ClimateProv[,c("Population","ClimaticMoistureDeficit")])) %>% 
  #arrange(-ClimaticMoistureDeficit)
RM_Tseason_tab = data.frame(unique(ClimateProv[,c("Population","Tseasonality")])) %>% 
  arrange(-Tseasonality)
#RM_Tmax_spring_tab = data.frame(unique(ClimateProv[,c("Population","Tmax_spring")])) %>% 
  #arrange(-Tmax_spring)
#I tested also Tmin_annual and Tmean_annual. It almost works with Tmean_annual (order close to thermicity index) but not Tmin.
#Thermicity index is indeed the right choice.

#Thermicity index takes into account annual temperature and minimum temperatures. Note that summer temperature is not explicitely in our models then.
#OakPalette = c("TUN_Tab" = "#D55E00", "SPA_LoAl" = "#E69F00", "TUN_Fer" = "#F0E442", "POR_QudSe" = "#D0D600", "SPA_Fue" = "#009E73", "ITA_Tus" = "#56B4E9", "FRA_MoeMa" = "#0072B2", "FRA_Pad" = "#CC79A7")#former palette according to Thermicity index
library(colorspace)

OakPalette <- sequential_hcl(
  n = 8,
  h = c(35, 120),     # orange to green
  c = c(80, 100),     # higher chroma for more saturation
  l = c(90, 35),      # bright warm -> darker cool
  power = 0.9         # slight boost for warm end
)

names(OakPalette) <- c("TUN_Tab","SPA_LoAl","TUN_Fer","POR_QudSe",
                       "SPA_Fue","ITA_Tus","FRA_MoeMa","FRA_Pad")

# Add a distinct green for the Cork oak climatic enveloppe
ClimEnvPalette <- c(OakPalette, "green")
names(ClimEnvPalette) <- c("TUN_Tab","SPA_LoAl","TUN_Fer","POR_QudSe",
                       "SPA_Fue","ITA_Tus","FRA_MoeMa","FRA_Pad",
                       "Cork oak climatic enveloppe")
```

#Map of Cork oak populations

```{r Map Cork oak populations}
Prov_coord = read.csv2("C:/Users/Marion/Documents/Thèse/2_Expérimentations/Oaks/Cork oak/Provenances coordinates.csv",dec=".") %>% 
  mutate(Population = case_when(provenance == "Moliets et Maa" ~ "FRA_MoeMa",
                                provenance == "Padouen" ~ "FRA_Pad",
                                provenance == "Fuencaliente" ~ "SPA_Fue",
                                provenance == "Tuscania" ~ "ITA_Tus",
                                provenance == "Quinta de Serra" ~ "POR_QudSe",
                                provenance == "Tabarka" ~ "TUN_Tab",
                                provenance == "Fernana" ~ "TUN_Fer",
                                provenance == "Los Alcornocales" ~ "SPA_LoAl",))
Distribution_area = read_sf("C:/Users/Marion/Documents/Thèse/3_Analyses & Modélisation/chorological_maps_dataset/Quercus suber/shapefiles/Quercus_suber_plg_clip.shp")
world <- ne_countries(scale = "medium", returnclass = "sf")

ggplot(data = world) +
  geom_sf(fill="white") +
  geom_sf(data = Distribution_area, fill="green", alpha = 0.3) +
  #geom_sf_text(aes(label = stringr::str_wrap(name, 1)), size = 3)+
  geom_point(data = Prov_coord, aes(x = longitude, y = latitude, fill = Population), size = 6, shape = 21) +
  scale_fill_manual(values = OakPalette, breaks=Prov_coord[,c("Population")])+
  geom_text_repel(data = Prov_coord, aes(x = longitude, y = latitude,label=as.character(Population)),hjust=-0.1,vjust=0.2,size=6)+
  coord_sf(xlim = c(-12, 25), ylim = c(33, 55), expand = FALSE)+
  theme_bw()+
  theme(axis.title=element_text(size=18),axis.text=element_text(size=16),legend.position ="none")+ xlab("Longitude")+ ylab("Latitude")+
  annotation_scale(location = "tr", width_hint = 0.5,text_cex = 1)+
  annotation_north_arrow(location = "tr", pad_x = unit(0.4, "in"), pad_y = unit(0.4, "in"), style = ggspatial::north_arrow_nautical(fill = c("grey40", "white"),line_col = "grey20",text_family = "ArcherPro Book"))
```

```{r Population localisation within the cork oak climatic enveloppe}
ClimEnveloppe <- read.csv("C:/Users/Marion/Documents/Thèse/2_Expérimentations/Oaks/cork oak/Building climatic raster on cork oak area/Climatic enveloppe Distribution area.csv",sep = ";") %>% 
  group_by(ID) %>% 
  summarise_all(mean) %>% 
  mutate(ID = "Cork oak climatic enveloppe")
colnames(ClimEnveloppe) = c("Provenance","Longitude","Latitude","Elevation","Year","Tseasonality","ClimaticMoistureDeficit","RM_ThermicityIndex")

ClimEnveloppeProv = ClimEnveloppe[,c("Provenance","Tseasonality","ClimaticMoistureDeficit","RM_ThermicityIndex")] %>% 
  rbind(ClimateProv[,c("Provenance","Tseasonality","ClimaticMoistureDeficit","RM_ThermicityIndex")]) %>% mutate(Population = case_when(Provenance == "Moliets et Maa" ~ "FRA_MoeMa",
                                Provenance == "Padouen" ~ "FRA_Pad",
                                Provenance == "Fuencaliente" ~ "SPA_Fue",
                                Provenance == "Tuscania" ~ "ITA_Tus",
                                Provenance == "Quinta de Serra" ~ "POR_QudSe",
                                Provenance == "Tabarka" ~ "TUN_Tab",
                                Provenance == "Fernana" ~ "TUN_Fer",
                                Provenance == "Los Alcornocales" ~ "SPA_LoAl",
                                Provenance == "Cork oak climatic enveloppe" ~ "Cork oak climatic enveloppe"))

ClimEnvPaletteExt <- ClimEnvPalette[ClimEnveloppeProv$Population]
ClimEnvSize = c("Tabarka" = 3, "Los Alcornocales" = 3, "Fernana" = 3, "Quinta de Serra" = 3, "Fuencaliente" = 3, "Tuscania" = 3, "Moliets et Maa" = 3, "Padouen" = 3,"Cork oak climatic enveloppe"=1.5)
ClimEnvSizeExt <- ClimEnvSize[ClimEnveloppeProv$Provenance]
ClimEnvAlpha = c("Tabarka" = 1, "Los Alcornocales" = 1, "Fernana" = 1, "Quinta de Serra" = 1, "Fuencaliente" = 1, "Tuscania" = 1, "Moliets et Maa" = 1, "Padouen" = 1,"Cork oak climatic enveloppe"=0.3)
ClimEnvAlphaExt <- ClimEnvAlpha[ClimEnveloppeProv$Provenance]

par3d(cex=3)
#plot3d(ClimEnveloppeProv$RM_ThermicityIndex, ClimEnveloppeProv$ClimaticMoistureDeficit, ClimEnveloppeProv$Tseasonality, col = ClimEnvPaletteExt, size = ClimEnvSizeExt, type = "s",alpha = ClimEnvAlphaExt,xlab = "R-M Thermicity Index", ylab = "Climatic Moisture Deficit", zlab = "Temperature Seasonality",box=F) 
plot3d(ClimEnveloppeProv$RM_ThermicityIndex, ClimEnveloppeProv$ClimaticMoistureDeficit, ClimEnveloppeProv$Tseasonality, col = ClimEnvPaletteExt, size = ClimEnvSizeExt, type = "s",alpha = ClimEnvAlphaExt,xlab = "", ylab = "", zlab = "",box=F) #no axis title

legend3d("topright", legend = names(ClimEnvPalette), col = ClimEnvPalette, pch = 16)

ggplot()+
  geom_point(aes(x=ClimaticMoistureDeficit,y=RM_ThermicityIndex,fill=Population),color="black",data=filter(ClimEnveloppeProv,Population == "Cork oak climatic enveloppe"),shape=21,size=5)+
  geom_point(aes(x=ClimaticMoistureDeficit,y=RM_ThermicityIndex,fill=Population),color="black",data=filter(ClimEnveloppeProv,Population != "Cork oak climatic enveloppe"),shape=21,size=10)+
  scale_fill_manual(values = ClimEnvPalette)+
  theme_bw()+ylab("Rivas-Martínez Thermicity Index")+xlab("Climatic Moisture Deficit")+
  theme(axis.text=element_text(size=30),axis.title=element_text(size=35),legend.title=element_text(size=18),legend.text=element_text(size=15))
ggplot()+
  geom_point(aes(x=Tseasonality,y=RM_ThermicityIndex,fill=Population),color="black",data=filter(ClimEnveloppeProv,Population == "Cork oak climatic enveloppe"),shape=21,size=5)+
  geom_point(aes(x=Tseasonality,y=RM_ThermicityIndex,fill=Population),color="black",data=filter(ClimEnveloppeProv,Population != "Cork oak climatic enveloppe"),shape=21,size=10)+
  scale_fill_manual(values = ClimEnvPalette)+
  theme_bw()+ylab("Rivas-Martínez Thermicity Index")+xlab("Temperature seasonality")+
  theme(axis.text=element_text(size=30),axis.title=element_text(size=35),legend.title=element_text(size=18),legend.text=element_text(size=15))
```

#Germination Data import

```{r Data import germination}
'%notin%' <- Negate('%in%')
#Germination data
BDDQuSub <- read_excel("C:/Users/Marion/Documents/Thèse/2_Expérimentations/Oaks/Cork oak/BDD_Quercus suber_2023.xlsx")
Germination = BDDQuSub %>% 
  dplyr::select(Provenance,Country,Longitude,Latitude,Collection_date,Id_mother,proxi_poids,DBH,Id_seed,Species,Date_semis,Day_temperature,Night_temperature,Germination,Germination_date,Germination_time) %>% 
  dplyr::mutate(Germination_time = as.numeric(Germination_time), Germination = as.numeric(Germination),
                Germination_date = as.numeric(Germination_date),
                Id_mother = as.character(paste(Provenance,"_",Id_mother),sep=""),
                Storage = as.numeric(Date_semis - Collection_date),
                DBH = as.numeric(DBH),
                Germination_date = format(as.Date(Germination_date, origin = "1899/12/30"),format = "%d/%m/%Y")) %>% #format de la date
  mutate(Population = case_when(Provenance == "Moliets et Maa" ~ "FRA_MoeMa",
                                Provenance == "Padouen" ~ "FRA_Pad",
                                Provenance == "Fuencaliente" ~ "SPA_Fue",
                                Provenance == "Tuscania" ~ "ITA_Tus",
                                Provenance == "Quinta de Serra" ~ "POR_QudSe",
                                Provenance == "Tabarka" ~ "TUN_Tab",
                                Provenance == "Fernana" ~ "TUN_Fer",
                                Provenance == "Los Alcornocales" ~ "SPA_LoAl")) %>% 
  filter(!is.na(Day_temperature)) %>% 
  group_by(Provenance,Day_temperature,Id_mother) %>% 
  dplyr::mutate(RoundAcornMass = round(mean(proxi_poids)))
```

#Germination indexes calculations

Which variables characterize germination ?
There are at least four conceptually distinct properties of the temporal pattern of germination that must be quantified in order to fully characterize it: delay, extent, average speed and variation in speed of germination (McNair et al, 2012) That is to say Time, percentage, and synchrony : germinability/germination rate/percentage, mean germination time, germination synchrony (MA Ranal and Santana, 2006).
- Germination percentage GP
- Mean Germination Time AT0 
- Germination Synchrony GS

A lot of germination indexes are not by acorns but by groups of provenance (calculated on 45 acorns) and/or chamber temperature (calculated on 15 acorns) and/or mother tree (calculated on 5 acorns).
Let's calculate them by provenance, temperature and mother.

```{r Germination indexes calculations}
GermIndexesRate = Germination[,c("Provenance","Population","Day_temperature","Night_temperature","Germination","Germination_time","proxi_poids","Id_mother")] %>%
  group_by(Population, Provenance, Day_temperature,Night_temperature,Id_mother) %>%
  dplyr::mutate(NbSeeds = n(),
                GP = GermPercent(germ.counts = Germination, total.seeds = mean(NbSeeds), partial = TRUE)/100,#germination percentage
                AM = mean(proxi_poids))#Mean acorn mass
GermIndexesTime = GermIndexesRate %>% 
  filter(Germination == 1) %>% 
  dplyr::mutate(NbSeedsGerm = n()) %>% 
  filter(NbSeedsGerm>2) %>% #remove mother with less than 2 seeds germinated
  group_by(Population, Day_temperature, Night_temperature,Id_mother, Germination_time) %>% 
  summarise(across(Germination, sum), across(c(proxi_poids,NbSeeds,GP,AM,NbSeedsGerm), mean)) %>%  
  group_by(Population, Day_temperature, Night_temperature,Id_mother) %>% 
  dplyr::mutate(AT0  = MeanGermTime(germ.counts = Germination, intervals = Germination_time, partial = TRUE),#Mean Germination Time
                GS = GermSynchrony(germ.counts = Germination, intervals = Germination_time, partial = TRUE),#Germination synchrony
                #Other indexes
                #CvGT = CVGermTime(germ.counts = Germination, intervals = Germination_time, partial = TRUE),#Coefficient of variation of the germination time
                #T50 = t50(germ.counts = Germination, intervals = Germination_time, partial = TRUE, method = c("coolbear", "farooq")),#median germination time
                #CUG = CUGerm(germ.counts = Germination, intervals = Germination_time, partial = TRUE),#Coefficient of uniformity of germination
                #FGT = FirstGermTime(germ.counts = Germination, intervals = Germination_time, partial = TRUE),#Time of first germination
                #LGT = LastGermTime(germ.counts = Germination, intervals = Germination_time, partial = TRUE),#Time of last germination
                #SGT = TimeSpreadGerm(germ.counts = Germination, intervals = Germination_time, partial = TRUE),#Time spread of germination
                #PGT = PeakGermTime(germ.counts = Germination, intervals = Germination_time, partial = TRUE),#Peak time of germination
                #GV = GermSpeed(germ.counts = Germination,intervals = Germination_time, partial = TRUE, percent = FALSE, total.seeds = NULL),#Speed of germination / Germination rate index 
                #GU = GermUncertainty(germ.counts = Germination,intervals = Germination_time, partial = TRUE),#Synchronization index or Uncertainty of the germination process
                #MGP = MeanGermRate(germ.counts = Germination,intervals = Germination_time, partial = TRUE),#Mean germination rate
                #CVeG = CVG(germ.counts = Germination,intervals = Germination_time, partial = TRUE)#Coefficient of velocity
                ) 

GerminationTraits = unique(GermIndexesRate[,c("Provenance", "Population", "Day_temperature", "Night_temperature", "Id_mother", "GP", "AM")]) %>% 
  left_join(unique(GermIndexesTime[,c("Population", "Day_temperature", "Night_temperature", "Id_mother", "AT0", "GS")]),by=c("Population", "Day_temperature", "Night_temperature", "Id_mother"))
```
Important note on the averaged T0 calculation:
For the PCA in the climatic characterization of populations, we calculated a group-level average T₀. For this, we used the MeanGermTime function from the R package germinationmetrics (Aravind et al., 2023). Although the function usually computes Mean Germination Time (AT0 ), because our observation intervals were irregular, it effectively returned the arithmetic mean of germination times across all seeds in a group (which is what we intended, since we already used synchrony to evaluate the spread of germination). For instance:
MeanGermTime(germ.counts = c(0,1,1,0,1,0,0,0), intervals = c(1,3,4,8,9,12,13,14), partial = TRUE)
[1] 5.333333
Warning message: In MeanGermTime 'intervals' are not uniform.
(3+4+9)/3
[1] 5.333333
Thus, in our dataset, what was previously labeled as “AT0 ” on the PCA plot (Figure 2) represents averaged T0, not the standard AT0  reflecting rate and spread of germination. To avoid confusion, we have relabeled it as AT0 (“Averaged T0”).


#Climate and Germination Data merge

```{r Merge climate germination}
#scale to reduce / avoid multicollinearity and to ensure that all variables contribute equally to the analysis and to prevent issues like numerical instability or disproportionate influence of certain features. 

GerminationScaled = Germination %>% mutate(proxi_poids = scale(proxi_poids), DBH = scale(DBH)) %>% left_join(ClimateProvScaled[,c("Provenance","ClimaticMoistureDeficit", "RM_ThermicityIndex", "DM_AridityIndex", "E_PluviometricQuotient", "Tseasonality", "SHM", "Pseasonality", "Pmean_summer", "Pmax_annual", "Tmean_annual", "Tmin_annual")],by="Provenance") %>% 
  left_join(ClimateProv[,c("Provenance","Tmax_spring")],by="Provenance")
GerminationScaled$Day_temperature = scale(as.numeric(GerminationScaled$Day_temperature))
GerminationScaled$Tmax_spring = scale(as.numeric(GerminationScaled$Tmax_spring))

Germination = Germination %>% left_join(ClimateProv[,c("Provenance","ClimaticMoistureDeficit", "Tmax_spring", "RM_ThermicityIndex", "DM_AridityIndex", "E_PluviometricQuotient", "Tseasonality", "SHM", "Pseasonality")],by="Provenance")

GerminationTraitsScaled = GerminationTraits[,c("Provenance","Population","Id_mother","GP","AT0","GS","Day_temperature","Night_temperature")] %>%  cbind(data.frame(AM = scale(GerminationTraits$AM))) %>%  left_join(ClimateProvScaled[,c("Provenance","ClimaticMoistureDeficit", "RM_ThermicityIndex", "DM_AridityIndex", "E_PluviometricQuotient", "Tseasonality", "SHM", "Pseasonality","Pmean_summer","Pmax_annual","Tmean_annual","Tmin_annual")],by="Provenance") %>% 
  left_join(ClimateProv[,c("Provenance","Tmax_spring")],by="Provenance")
GerminationTraitsScaled = GerminationTraitsScaled %>% ungroup() %>% mutate(Id=1:nrow(GerminationTraitsScaled))
GerminationTraitsScaled$Day_temperature = scale(as.numeric(GerminationTraitsScaled$Day_temperature))
GerminationTraitsScaled$Tmax_spring = scale(as.numeric(GerminationTraitsScaled$Tmax_spring))

BothScaled = Germination[,c("Provenance","Id_mother","Id_seed","Germination","Night_temperature")] %>% cbind(data.frame(proxi_poids = scale(Germination$proxi_poids),DBH = scale(Germination$DBH),Germination_time = scale(Germination$Germination_time),RoundAcornMass = scale(Germination$RoundAcornMass))) %>%
  left_join(GerminationTraitsScaled,by=c("Provenance","Id_mother","Night_temperature"))
```

#Characterization of populations with climatic variables 

```{r  ACP ALL}
ACP = PCA(filter(BothScaled[,c("Population","GP","AT0","GS","Tmax_spring","ClimaticMoistureDeficit","RM_ThermicityIndex","Tseasonality")] %>% rename("TSp" = "Tmax_spring", "CMD" = "ClimaticMoistureDeficit", "RMTh" = "RM_ThermicityIndex", "TSn" = "Tseasonality"),
                 !is.na(AT0 )), scale.unit =TRUE, ncp = 6, graph = F, quali.sup = c(1))

ACPVarPalette = c("Trait" = "black", "Water stress" = "#004488","Spring temperature" = "#CD3E4E","Temperature seasonality" = "#882255","Temperature (minimal, \n annual and spring)" = "#CD3E4E")

BothScaled$Population <- factor(BothScaled$Population , levels=c("TUN_Tab", "SPA_LoAl", "TUN_Fer", "POR_QudSe", "SPA_Fue", "ITA_Tus", "FRA_MoeMa", "FRA_Pad"))#order population level
#grid.arrange(
fviz_pca_biplot(ACP, axes = c(1, 2), pointshape = 21, pointsize = 3, col.ind = "black", fill.ind = filter(BothScaled[,c("Population","GP","AT0","GS","Tmax_spring","ClimaticMoistureDeficit","RM_ThermicityIndex","Tseasonality")] %>% rename("TSp" = "Tmax_spring", "CMD" = "ClimaticMoistureDeficit", "RMTh" = "RM_ThermicityIndex", "TSn" = "Tseasonality"),
                 !is.na(AT0 ))$Population,  palette = OakPalette, fill.var = "white",label = c("var","quali.sup"), habillage = "none",legend.title = list(color = "Variable type", fill = "Population"),repel = T, col.var = factor(c("Trait","Trait","Trait","Spring temperature","Water stress","Temperature (minimal, \n annual and spring)","Temperature seasonality")),labelsize = 8)+
  scale_color_manual(values = ACPVarPalette)+theme_bw()+
  theme(legend.spacing.y = unit(0.3, 'cm'),text = element_text(size = 20),axis.title = element_text(size = 15),axis.text = element_text(size = 10)) + guides(fill = guide_legend(byrow = TRUE),color = guide_legend(byrow = TRUE))#,
#fviz_pca_biplot(ACP, axes = c(1, 3), pointshape = 21, col.ind = "black", fill.ind = filter(BothScaled[,c("Population","GP","Germination_time","AT0","GS","CvGT","Tmax_spring","ClimaticMoistureDeficit","RM_ThermicityIndex","Tseasonality")],!is.na(Germination_time) & !is.na(AT0 ))$Population,  palette = OakPalette, fill.var = "white",label = c("var","quali.sup"), habillage = "none",legend.title = list(color = "Type", fill = "Population"),repel = T,col.var = factor(c("Trait","Trait","Trait","Trait","Trait","Spring temperature","Water stress","Cold intensity and \n annual temperature","Temperature seasonality")))+ 
#scale_color_manual(values = ACPVarPalette),
#nrow = 1)
fviz_eig(ACP, addlabels = TRUE, ylim = c(0, 50))#Graphe des éboulis
```

#General models and specific Population models on each Trait

1) General models on each trait: 
  Trait ~ Population + Day_temperature + Population:Day_temperature
2) Specific models on each traits:
  GP / G(0,1) ~ Population climatic variables
GS ~ Population climatic variables
AT0  / GT ~ Population climatic variables + Day_temperature + Population climatic variables:Day_temperature

```{r  Models assumptions checking}
#The interpretation of conventional residuals for generalized linear (mixed) and other hierarchical statistical models is often problematic. 
#(usual rules for interpreting residuals plot do not apply - This is the appearance you expect of such a plot when the dependent variable is discrete. https://stats.stackexchange.com/questions/25068/interpreting-plot-of-residuals-vs-fitted-values-from-poisson-regression)
# --> DHARMa package https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html
ResidualsOutput = simulateResiduals(fittedModel = Model, plot = F,quantreg=T)#first calculate residuals
par(
  mfrow = c(1, 1),  # 2x2 layout
  mar = c(6, 6, 3, 2),  # more space
  las = 1,          # horizontal labels
  cex.axis = 1.2,   # slightly larger axis text
  cex.lab = 1.3,    # slightly larger axis titles
  cex.main = 1.2    # larger titles
)
plotQQunif(ResidualsOutput)
plotResiduals(ResidualsOutput)
plot(ResidualsOutput)#Scaled residuals plot
vif(Model)
plotresid(Model)#independance and variance (and normality if normal based model)
#you can also use it to plot residuals against a specific other predictors (highly recommend).
plotResiduals(ResidualsOutput, GerminationT$RM_ThermicityIndex)
testDispersion(ResidualsOutput)
#Visual global check of model assumptions (some are classic residuals graphs that must not be interpreted)
#windows(width = 12, height = 10)
check_model(Model) ; vif(Model)
```

```{r  Stepwise selection}
options(na.action = "na.fail") # Required for dredge to run
ModelSelection <- dredge(Model, rank = AIC, evaluate = T, extra = c(R2m = function(x) r.squaredGLMM(x)[1,1], R2c = function(x) r.squaredGLMM(x)[1,2]))
#or
ModelSelection <- dredge(Model, rank = AIC, evaluate = T)
options(na.action = "na.omit")
get.models(ModelSelection, subset = 1)[[1]]
summary(model.avg(ModelSelection, subset = delta <= 1))
```

##Germination probability (Cox models)

```{r Germination Curves on raw data}
#Dataframe nombre de graines
NbSeeds = Germination[,c("Population","Day_temperature","Id_seed")] %>% 
  dplyr::mutate(compteur = 1) %>% 
  group_by(Day_temperature, Population) %>%
  dplyr::summarise(Nb_SeedsProv = sum(compteur)) %>%
  group_by(Day_temperature) %>% 
  dplyr::mutate(Nb_Seeds = sum(Nb_SeedsProv))
#Graphe cumulatif des germinations au fur et à mesure du temps
CumulGerminDay = filter(Germination, !is.na(Germination_date)) %>%
  left_join(NbSeeds[,c("Day_temperature", "Population", "Nb_SeedsProv")], by = c( "Day_temperature", "Population")) %>% 
  left_join(unique(NbSeeds[,c("Day_temperature", "Nb_Seeds")]), by = c("Day_temperature")) %>%
  group_by(Day_temperature, Germination_date, Germination_time, Nb_SeedsProv, Population) %>% 
  dplyr::summarise(Nb_germ = sum(Germination)) %>% 
  arrange(Day_temperature, Population,Germination_time) %>% 
  group_by(Day_temperature,Population) %>% 
  dplyr::mutate(Nb_germCumul = cumsum(Nb_germ),
                Perc_germCumul = Nb_germCumul/Nb_SeedsProv) %>% 
  group_by(Day_temperature, Germination_date, Germination_time,Population) %>% #Pour avoir un résultat par jour et éviter les doublons
  dplyr::summarise(Nb_germCumul = max(Nb_germCumul),Perc_germCumul = max(Perc_germCumul)) %>% 
  arrange(Day_temperature, Population,Germination_time)
#Courbes de germination données brutes
ggplot(CumulGerminDay, aes(Germination_time, Perc_germCumul))+
  geom_point()+ geom_smooth(span = 1, aes(color = Population, fill = Population),size=1.5) + ylim(0,1) + xlim(0,50)+
  scale_color_manual(values = OakPalette,breaks=CMD_tab[,c("Population")])+
  scale_fill_manual(values = OakPalette,breaks=CMD_tab[,c("Population")])+
  xlab("Germination time")+ ylab("Germination rate")+
  theme_bw()+
  theme(plot.title = element_markdown(box.color = "black", color = "white", fill = "darkgreen", hjust = 0.5, linewidth = 0.5, r = unit(0, "mm"), linetype = 1, padding = unit(2, "mm")),axis.title = element_text(size = 30),axis.text = element_text(size = 20),strip.text = element_text(size = 20))+ guides(fill="none")+
  facet_wrap(~paste(Day_temperature, " °C"),nrow = 3)
ggplot(CumulGerminDay, aes(Germination_time, Perc_germCumul))+
  geom_point()+ geom_smooth(span = 1, aes(color = as.factor(Day_temperature), fill = as.factor(Day_temperature))) + ylim(0,1) +
  scale_fill_manual(values = c("15" = "#3399FF", "20" = "#009900", "25" = "#CC3333"))+
  scale_color_manual(values = c("15" = "#3399FF", "20" = "#009900", "25" = "#CC3333"))+
  xlab("Germination time")+ ylab("Germination rate")+
  theme_bw()+
  theme(plot.title = element_markdown(box.color = "black", color = "white", fill = "darkgreen", hjust = 0.5, linewidth = 0.5, r = unit(0, "mm"), linetype = 1, padding = unit(2, "mm")))+ guides(fill="none")+
  facet_wrap(~ Population, nrow = 2)
```

```{r General cox model run and total predictions}
#Coxfit = coxph(Surv(Germination_time, Germination) ~ Population + Population:Day_temperature + Day_temperature + proxi_poids + proxi_poids:Population, data=Germination, robust = TRUE)#Tmax_spring not necessary (AIC remains the same). Mother tree increase the AIC and RoundAcornmass decrease it so we remove Mother tree and keep RoundAcornMass (Mean acorn mass per mother)
#Coxfit = coxph(Surv(Germination_time, Germination) ~ Population + Population:as.factor(Day_temperature) + as.factor(Day_temperature) + proxi_poids + proxi_poids:Population, data=Germination, robust = TRUE)
#Acorn mass removed after GLMMs stability tests: if the effect is not stable for Germination time and Germination status, we have to remove it too from Cox models, which are using both germination time and germination status. 
Coxfit = coxph(Surv(Germination_time, Germination) ~ Population + Population:as.factor(Day_temperature) + as.factor(Day_temperature), data=Germination, robust = TRUE)
#Assumptions/hypothesis checking
test.ph = cox.zph(Coxfit)#checking proportional Hazards assumption #not assumed on the entire date, but assumed on a sample --> ok ?
summary(Coxfit)
Ano = data.frame(car::Anova(Coxfit))
Covar_val = unique(Germination[,c("Population","Day_temperature","RoundAcornMass","proxi_poids")])
Survivalfit = survfit(Coxfit, newdata = Covar_val)

#Predictions per population, chamber and mother/RoundAcornMass
Predict_extractTOT = survfit2df(Survivalfit) %>% 
  left_join(cbind(Covar_val,row.names(Covar_val)) %>% rename("strata"="...5"),by = "strata") %>% 
  mutate(PH = cox.zph(Coxfit)[["table"]][6,3],
         surv = 1 - surv,
         lower = 1 - lower,
         upper = 1 - upper,
         Chisq_Prov = c(round(Ano[1,1],2)),
         Chisq_dayT = c(round(Ano[2,1],2)),
         Chisq_AM = c(round(Ano[3,1],2)),
         pval_Prov = Ano[1,3],
         pval_dayT = Ano[2,3],
         pval_AM = Ano[3,3])
```

```{r Cox model Predictions}
#Population within day_t
Predict_extractProv = Predict_extractTOT %>% group_by(Population,Day_temperature,time) %>% summarise_at(vars(surv, lower,upper,pval_Prov,Chisq_Prov), funs(mean)) %>% 
  dplyr::mutate(pval_Prov = ifelse(pval_Prov<0.001, "<0.001",as.character(round(pval_Prov,3)))) 

#Pairwise comparisons
EMMP = emmeans(Coxfit, ~ Population | Day_temperature, mode="link")
CompMultP = back.emmeans(EMMP,n=120)  ; colnames(CompMultP) = c("Tosep","SE_inf","EMMean","SE_sup")  ; CompMultP = separate(CompMultP, col = "Tosep", into = c("Population","Day_temperature"), sep = ":", extra = "merge") %>% group_by(Day_temperature) %>% arrange(EMMean)
CompMultP = CompMultP %>% left_join(cld(EMMP, adjust = "tukey", Letters = LETTERS)[,c("Day_temperature","Population",".group")] %>% mutate(Day_temperature = as.character(Day_temperature)), by = c("Day_temperature","Population")) %>% group_by(Day_temperature) %>% mutate(Population = factor(Population, levels = unique(Population)))
ggplot(data = CompMultP,aes(x=reorder(Day_temperature, EMMean),y=EMMean, group = reorder_within(Population, EMMean, Day_temperature), color=Population))+
  geom_point(position=position_dodge(width=0.9), size = 3)+
  geom_errorbar(aes(ymin = SE_inf, ymax = SE_sup), position=position_dodge(width=0.9),size = 2)+
  geom_text(aes(label = CompMultP$.group,GPoup = reorder_within(Population, EMMean, Day_temperature), color=Population,y=SE_inf-0.05), position=position_dodge(width=0.9), size = 6) +
  scale_color_manual(name = "Population",values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")]) +
  xlab("Day_temperature")+ylab("Germination probability")+
  theme_bw() + theme(axis.text=element_text(size=25),axis.title = element_text(size = 30), legend.text = element_text(size=20),legend.title = element_text(size=25),legend.key.size = unit(1.5, 'cm'))

ggplot()+
  geom_line(data = Predict_extractProv, size = 1.5,aes(x = time, y = surv, color = Population))+#geom_smooth(size = 1, se=F)+
  scale_color_manual(values = OakPalette,breaks=RM_ThermicityIndex_tab[,c("Population")])+
  geom_hline(yintercept=0.5, linetype="dashed", color = "black", size = 1)+
  geom_ribbon(data = Predict_extractProv, aes(ymin = lower, ymax = upper, fill = Population,x = time, y = surv), alpha = 0.2,colour = NA)+
  scale_fill_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])+
  xlab("Time")+ylab("Germination probability (percentage)")+
  theme_bw() +
  geom_text_repel(aes(label = Group,x = time, y = surv, color = Population), size = 7, show.legend = FALSE, fontface = "bold",
                  data = CompMultP[,c("Population","Day_temperature",".group")] %>% 
                    mutate(Day_temperature = as.numeric(Day_temperature)) %>% 
                    left_join(filter(Predict_extractProv[,c("Population","Day_temperature","time","surv")],surv<0.95 & surv > 0.80), by=c("Population","Day_temperature")) %>% 
                    distinct(Population, Day_temperature, .keep_all = TRUE) %>% 
                    mutate(Group = as.character(.group)))+
  theme(axis.title = element_text(size = 35),axis.text = element_text(size = 30),strip.text = element_text(size = 30),legend.title=element_text(size=18),legend.text=element_text(size=15))+
  facet_wrap(~ Day_temperature, nrow = 3)

#day_t within Population
Predict_extractDayT = Predict_extractTOT %>% group_by(Population,Day_temperature,time) %>% summarise_at(vars(surv, lower,upper,pval_dayT,Chisq_dayT), funs(mean)) %>% 
  dplyr::mutate(pval_dayT = ifelse(pval_dayT<0.001, "<0.001",as.character(round(pval_dayT,3))))

#Pairwise comparisons
EMMP = emmeans(Coxfit, ~ Day_temperature | Population, mode="link")
CompMultP = back.emmeans(EMMP,n=120)  ; colnames(CompMultP) = c("Tosep","SE_inf","EMMean","SE_sup")  ; CompMultP = separate(CompMultP, col = "Tosep", into = c("Day_temperature","Population"), sep = ":", extra = "merge") %>% group_by(Population) %>% arrange(EMMean)
CompMultP = CompMultP %>% left_join(cld(EMMP, adjust = "tukey", Letters = LETTERS)[,c("Population","Day_temperature",".group")] %>% mutate(Day_temperature = as.character(Day_temperature)), by = c("Population","Day_temperature")) %>% group_by(Population) %>% mutate(Day_temperature = factor(Day_temperature, levels = unique(Day_temperature)))
ggplot(data = CompMultP,aes(x=reorder(Population, EMMean),y=EMMean, group = reorder_within(Day_temperature, EMMean, Population), color=Day_temperature))+
  geom_point(position=position_dodge(width=0.9), size = 3)+
  geom_errorbar(aes(ymin = SE_inf, ymax = SE_sup), position=position_dodge(width=0.9),size = 2)+
  geom_text(aes(label = CompMultP$.group,GPoup = reorder_within(Day_temperature, EMMean, Population), color=Day_temperature,y=SE_inf-0.05), position=position_dodge(width=0.9), size = 6) +
  #scale_color_manual(name = "Day_temperature",values = OakPalette, breaks=TmaxSpring_tab[,c("Day_temperature")]) +
  xlab("Population")+ylab("Germination probability")+
  theme_bw() + theme(axis.text=element_text(size=25),axis.title = element_text(size = 30), legend.text = element_text(size=20),legend.title = element_text(size=25),legend.key.size = unit(1.5, 'cm'))

ggplot(data = Predict_extractDayT, aes(x = time, y = surv, color = as.factor(Day_temperature)))+
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = as.factor(Day_temperature)), alpha = 0.2,colour = NA)+
  scale_fill_manual(values = c("15" = "#0072B2", "20" = "#F0E442", "25" = "#D55E00"),name = "Chamber temperature")+
  geom_line(size = 1.5)+#geom_smooth(size = 1, se=F)+
  scale_color_manual(values = c("15" = "#0072B2", "20" = "#F0E442", "25" = "#D55E00"),name = "Chamber temperature")+
  geom_hline(yintercept=0.5, linetype="dashed", color = "black", size = 1)+
  theme(legend.position = "bottom")+
  xlab("Time")+ylab("Germination probability (percentage)")+
  theme_bw()+
  #geom_text(aes(x = 40, y = 0.25, label= paste("p = ", pval_dayT, "\n", "Chisq"," = ", Chisq_dayT)),color = "black", check_overlap = TRUE)+
  geom_text_repel(aes(label = Group,x = time, y = surv, color = as.factor(Day_temperature)), size = 7, show.legend = FALSE, fontface = "bold",
                  data = CompMultP[,c("Population","Day_temperature",".group")] %>% 
                    left_join(filter(Predict_extractDayT[,c("Population","Day_temperature","time","surv")],surv<0.95 & surv > 0.80)%>% 
                                mutate(Day_temperature = as.character(Day_temperature)), by=c("Population","Day_temperature")) %>% 
                    distinct(Population, Day_temperature, .keep_all = TRUE) %>% 
                    mutate(Group = as.character(.group)))+
  theme(axis.title = element_text(size = 35),axis.text = element_text(size = 30),strip.text = element_text(size = 30), legend.title=element_text(size=18),legend.text=element_text(size=15))+
  facet_wrap2( ~ factor(Population, levels=c("TUN_Tab", "SPA_LoAl", "TUN_Fer", "POR_QudSe", "SPA_Fue", "ITA_Tus", "FRA_MoeMa", "FRA_Pad")), ncol=4, nrow = 2,strip = strip_themed(background_x = elem_list_rect(fill = OakPalette)))

ggplot(data = Predict_extractDayT, aes(x = time, y = surv, color = as.factor(Day_temperature)))+
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = as.factor(Day_temperature)), alpha = 0.2,colour = NA)+
  scale_fill_manual(values = c("15" = "#0072B2", "20" = "#F0E442", "25" = "#D55E00"),name = "Chamber temperature")+
  geom_line(size = 1.5)+#geom_smooth(size = 1, se=F)+
  scale_color_manual(values = c("15" = "#0072B2", "20" = "#F0E442", "25" = "#D55E00"),name = "Chamber temperature")+
  geom_hline(yintercept=0.5, linetype="dashed", color = "black", size = 1)+
  theme(legend.position = "bottom")+
  xlab("Time")+ylab("Germination probability (percentage)")+
  theme_bw()+
  #geom_text(aes(x = 40, y = 0.25, label= paste("p = ", pval_dayT, "\n", "Chisq"," = ", Chisq_dayT)),color = "black", check_overlap = TRUE)+
  geom_text_repel(aes(label = Group,x = time, y = surv, color = as.factor(Day_temperature)), size = 7, fontface = "bold",
                  data = CompMultP[,c("Population","Day_temperature",".group")] %>% 
                    left_join(filter(Predict_extractDayT[,c("Population","Day_temperature","time","surv")],surv<0.95 & surv > 0.80)%>% 
                                mutate(Day_temperature = as.character(Day_temperature)), by=c("Population","Day_temperature")) %>% 
                    distinct(Population, Day_temperature, .keep_all = TRUE) %>% 
                    mutate(Group = as.character(.group)))+
  theme(axis.title = element_text(size = 35),axis.text = element_text(size = 30),strip.text = element_text(size = 30), legend.title=element_text(size=18),legend.text=element_text(size=15))+
  facet_wrap2( ~ factor(Population, levels=c("Fuencaliente", "Tuscania", "Fernana", "Padouen", "Tabarka", "Moliets et Maa", "Los Alcornocales", "Quinta de Serra")), ncol=4, nrow = 2,strip = strip_themed(background_x = elem_list_rect(fill = OakPalette[c("Fuencaliente", "Tuscania", "Fernana", "Padouen", "Tabarka", "Moliets et Maa", "Los Alcornocales", "Quinta de Serra")])))
```

```{r General cox model Output tables}
CoxGanova_table = rbind(c("Germination probability","","",""),Ano %>% rownames_to_column(., "Covariate") %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "RoundAcornMass", "Acorn Mass"))) %>% rename(p = "Pr..Chisq.") %>%  
                          mutate_at(vars(c("Chisq")), function(x) round(as.numeric(x), 2)) %>% select(Covariate,Chisq,Df,p))
CoxGsummary_table = rbind(c("Germination probability","","","",""),as.data.frame(summary(Coxfit)[["coefficients"]]) %>% rownames_to_column(., "Covariate") %>%  mutate_all(funs(str_replace(., "Population", ""))) %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "RoundAcornMass", "Acorn Mass"))) %>% 
                            mutate_at(vars(c("coef","exp(coef)","se(coef)","robust se","z")), function(x) round(as.numeric(x), 2)) %>% rename(p = "Pr(>|z|)","Test value" = "z","Std. Error" = "robust se", "Estimate" ="exp(coef)") %>% select(Covariate, Estimate, "Std. Error", "Test value", p))
GPH_table = as.data.frame(cox.zph(Coxfit)[["table"]]) %>% rownames_to_column(., "Covariate") %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "RoundAcornMass", "Acorn Mass"))) %>%  
  mutate_at(vars(c("chisq")), function(x) round(as.numeric(x), 2))
CoxGPerf_table = model_performance(Coxfit) %>% mutate(Model = "General GProba model")

Times = Predict_extractTOT %>% 
  group_by(Population, Day_temperature) %>% 
  summarise(IniTime = as.numeric(approx(surv, time, xout=0.03)),
            HalfTime = as.numeric(approx(surv, time, xout=0.5)),
            EndTime = as.numeric(approx(surv, time, xout=0.97))) %>% #0.97 et pas 1 pour ne pas avoir de NA
  filter(HalfTime != 0.5 & EndTime != 0.97 & IniTime != 0.01) %>% mutate(HalfTime = round(HalfTime,digits = 1), EndTime = round(EndTime,digits = 1),IniTime = round(IniTime,digits = 1)) %>% 
  mutate(TotalTime = EndTime - IniTime)

Times = Predict_extractDayT %>% 
  group_by(Population, Day_temperature) %>% 
  summarise(HalfTime = as.numeric(approx(surv, time, xout=0.5))) %>% 
  filter(HalfTime != 0.5) %>% mutate(HalfTime = round(HalfTime,digits = 1))

times_table = as.data.frame(Times[,c("Population","Day_temperature","HalfTime")] %>% filter(Day_temperature == 15) %>% 
                              left_join(Times[,c("Population","Day_temperature","HalfTime")] %>% filter(Day_temperature == 20),by = "Population") %>% 
                              left_join(Times[,c("Population","Day_temperature","HalfTime")] %>% filter(Day_temperature == 25),by = "Population")) %>% 
  rowwise() %>% 
  mutate(meanHalf = mean(c(HalfTime.x,HalfTime.y,HalfTime)),
         sdHalf = sd(c(HalfTime.x,HalfTime.y,HalfTime)))

times_table = times_table[order(times_table$meanHalf,decreasing=FALSE),] 
```

```{r Specific cox model run and total predictions}
Coxfit = coxph(Surv(Germination_time, Germination) ~ ClimaticMoistureDeficit + Tmax_spring + Tseasonality + RM_ThermicityIndex + Day_temperature +  ClimaticMoistureDeficit:Day_temperature + Tmax_spring:Day_temperature + RM_ThermicityIndex:Day_temperature + Tseasonality:Day_temperature, data=GerminationScaled, robust = TRUE)
test.ph = cox.zph(Coxfit)#checking proportional Hazards assumption #not assumed on the entire date, but assumed on a sample --> ok ?
ggcoxzph(test.ph)#checking proportional Hazards assumption graphic
plot(test.ph)
ggcoxdiagnostics(CoxfitT15, type = "dfbeta",# checking influential observations
                 linear.predictions = FALSE, ggtheme = theme_bw())
ggcoxfunctional(Surv(Germination_time, Germination) ~ Population, data=filter(Germination, Day_temperature == 15))#checking linearity
summary(Coxfit)
Ano = data.frame(car::Anova(Coxfit))
#rownames(Ano) <- NULL
Covar_val = unique(Germination[,c("ClimaticMoistureDeficit","Tmax_spring","Tseasonality","RM_ThermicityIndex","Day_temperature")])
Survivalfit = survfit(Coxfit, newdata = Covar_val)

#Predictions per population, chamber and mother/RoundAcornMass
Predict_extractTOT = survfit2df(Survivalfit) %>% 
  left_join(cbind(Covar_val,row.names(Covar_val)) %>% rename("strata"="...6"),by = "strata") %>% 
  mutate(PH = cox.zph(Coxfit)[["table"]][5,3],
         surv = 1 - surv,
         lower = 1 - lower,
         upper = 1 - upper,
         Chisq_CMD = c(round(Ano[1,2],2)),
         Chisq_Tspring = c(round(Ano[2,2],2)),
         Chisq_Tseason = c(round(Ano[3,2],2)),
         Chisq_ThermI = c(round(Ano[4,2],2)),
         Chisq_CMD = Ano[1,3],
         Chisq_Tspring = Ano[2,3],
         Chisq_Tseason = Ano[3,3],
         Chisq_ThermI = Ano[4,3])
```

```{r Specific cox model Output tables}
CoxSanova_table = rbind(c("Germination probability","","",""),Ano %>% rownames_to_column(., "Covariate") %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "RoundAcornMass", "Acorn Mass"))) %>% rename(p = "Pr..Chisq.") %>%  
                          mutate_at(vars(c("Chisq")), function(x) round(as.numeric(x), 2)) %>% select(Covariate,Chisq,Df,p))
CoxSsummary_table = rbind(c("Germination probability","","","",""),as.data.frame(summary(Coxfit)[["coefficients"]]) %>% rownames_to_column(., "Covariate") %>%  mutate_all(funs(str_replace(., "Population", ""))) %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "RoundAcornMass", "Acorn Mass"))) %>% 
                            mutate_at(vars(c("coef","exp(coef)","se(coef)","robust se","z")), function(x) round(as.numeric(x), 2)) %>% rename(p = "Pr(>|z|)","Test value" = "z","Std. Error" = "robust se", "Estimate" ="exp(coef)") %>% select(Covariate, Estimate, "Std. Error", "Test value", p))
SPH_table = as.data.frame(cox.zph(Coxfit)[["table"]]) %>% rownames_to_column(., "Covariate") %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "RoundAcornMass", "Acorn Mass"))) %>%  
  mutate_at(vars(c("chisq")), function(x) round(as.numeric(x), 2))
CoxSPerf_table = model_performance(Coxfit) %>% mutate(Model = "Specific GProba model")

Times = Predict_extractTOT %>% 
  group_by(Population, Day_temperature) %>% 
  summarise(IniTime = as.numeric(approx(surv, time, xout=0.03)),
            HalfTime = as.numeric(approx(surv, time, xout=0.5)),
            EndTime = as.numeric(approx(surv, time, xout=0.97))) %>% #0.97 et pas 1 pour ne pas avoir de NA
  filter(HalfTime != 0.5 & EndTime != 0.97 & IniTime != 0.01) %>% mutate(HalfTime = round(HalfTime,digits = 1), EndTime = round(EndTime,digits = 1),IniTime = round(IniTime,digits = 1)) %>% 
  mutate(TotalTime = EndTime - IniTime)
times_table = as.data.frame(Times[,c("Population","Day_temperature","HalfTime","TotalTime")] %>% filter(Day_temperature == 15) %>% 
                              left_join(Times[,c("Population","Day_temperature","HalfTime","TotalTime")] %>% filter(Day_temperature == 20),by = "Population") %>% 
                              left_join(Times[,c("Population","Day_temperature","HalfTime","TotalTime")] %>% filter(Day_temperature == 25),by = "Population"))

times_table = times_table[order(times_table$HalfTime.y,decreasing=FALSE),]
```

##G(0,1) / GP

Response variable: Germination -> binomial data
-> Generalized logistic mixed-effects models (glmer)

```{r  G(0,1) general model}
#Model1 = glm(Germination ~  Day_temperature + Population + Day_temperature:Population + proxi_poids + proxi_poids:Population, family = binomial(link = "logit"), data = GerminationScaled)
#Model2 = glmer(Germination ~  Day_temperature + Population + Day_temperature:Population + proxi_poids + proxi_poids:Population + (1 | Id_mother), family = binomial(link = "logit"), data = GerminationScaled, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))
#AIC(Model1,Model2)#AIC lower with random effect

Model = glmer(Germination ~  Day_temperature + Population + Day_temperature:Population + proxi_poids + proxi_poids:Population + (1 | Id_mother), family = binomial(link = "logit"), data = GerminationScaled, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

model_performance(Model) ; r.squaredGLMM(Model)#Variance expliqué (variance marginale des effets fixes et variance conditionnelle fixe + random)
Anova(Model)
summary(Model)

#Cross validation
Germination_training = GerminationScaled[sample(nrow(GerminationScaled), 400), ]#training data set
Germination_validate = filter(GerminationScaled, Id_seed %notin% Germination_training$Id_seed) %>% group_by(Population, Day_temperature) %>% mutate(Germination_rate = sum(Germination)/n())
ModelCV = glmer(Germination ~  Day_temperature + Population + Day_temperature:Population + proxi_poids + proxi_poids:Population + (1 | Id_mother), family = binomial(link = "logit"), data = Germination_training, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))#training of the Model on the training dataset
Predictions = predict(ModelCV, Germination_validate, re.form=NA, type="response")
cor.test(Predictions, Germination_validate$Germination_rate)
scatter_plot(obs = Germination_validate$Germination_rate, pred = Predictions) + coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))#print obs VS pred on a graphe 1:1

#Output tables
GGanova_table = rbind(c("G(0,1)","","",""),as.data.frame(Anova(Model)) %>% rownames_to_column(., "Covariate") %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "Pr(>Chisq)") %>% 
                        mutate_at(vars(c("Chisq")), function(x) round(as.numeric(x), 2)))
GGsummary_table = rbind(c("G(0,1)","","","",""),as.data.frame(summary(Model)[["coefficients"]]) %>% rownames_to_column(., "Covariate") %>%  mutate_all(funs(str_replace(., "Population", ""))) %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "Pr(>|z|)") %>% 
                          mutate_at(vars(c("Estimate","Std. Error","z value")), function(x) round(as.numeric(x), 2)) %>% rename("Test value" = "z value"))
GGPerf_table = model_performance(Model) %>% mutate(PearsonsCorrelation_obsVSpred = cor.test(Predictions, Germination_validate$Germination_rate)[["estimate"]][["cor"]],Model = "General GP model")
```

```{r  G(0,1) specific Population model}
Model = glmer(Germination ~  ClimaticMoistureDeficit + Tmax_spring + Tseasonality + RM_ThermicityIndex + (1 | Id_mother), family = binomial(link = "logit"), data = GerminationScaled)
#Stepwise selection:doesn't change the result

model_performance(Model) ; r.squaredGLMM(Model)#Variance expliqué (variance marginale des effets fixes et variance conditionnelle fixe + random)
Anova(Model)
summary(Model)

#Cross validation
Germination_training = GerminationScaled[sample(nrow(GerminationScaled), 400), ]#training data set
Germination_validate = filter(GerminationScaled, Id_seed %notin% Germination_training$Id_seed) %>% group_by(Population, Day_temperature) %>% mutate(Germination_rate = sum(Germination)/n())
ModelCV = glm(Germination ~  ClimaticMoistureDeficit + Tmax_spring + Tseasonality + RM_ThermicityIndex, family = binomial(link = "logit"), data = Germination_training)#training of the Model on the training dataset
Predictions = predict(ModelCV, Germination_validate, re.form=NA, type="response")
cor.test(Predictions, Germination_validate$Germination_rate)
scatter_plot(obs = Germination_validate$Germination_rate, pred = Predictions) + coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))#print obs VS pred on a graphe 1:1

#Output tables
GSanova_table = rbind(c("G(0,1)","","",""),as.data.frame(Anova(Model)) %>% rownames_to_column(., "Covariate") %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "Pr(>Chisq)") %>%  
                        mutate_at(vars(c("Chisq")), function(x) round(as.numeric(x), 2)))
GSsummary_table = rbind(c("G(0,1)","","","",""),as.data.frame(summary(Model)[["coefficients"]]) %>% rownames_to_column(., "Covariate") %>%  mutate_all(funs(str_replace(., "Population", ""))) %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "Pr(>|z|)") %>% 
                          mutate_at(vars(c("Estimate","Std. Error","z value")), function(x) round(as.numeric(x), 2)) %>% rename("Test value" = "z value"))
GSPerf_table = model_performance(Model) %>% mutate(PearsonsCorrelation_obsVSpred = cor.test(Predictions, Germination_validate$Germination_rate)[["estimate"]][["cor"]],Model = "Specific GP model")
```

Index
```{r  GP general model}
ggplot(data = GerminationTraits)+geom_histogram(aes(GR))#quantitative continuous between 0 an 1 (proportion): beta regression
ggplot(data = GerminationTraits)+geom_histogram(aes(p.beta(GerminationTraits$GP)))#put it in ]0; 1[ instead of [0; 1] 

Model = betareg(p.beta(GerminationTraitsScaled$GP) ~ Day_temperature + Population + Day_temperature:Population + AM + AM:Population, data = GerminationTraitsScaled)
#Model = glmmTMB(p.beta(GR) ~ Day_temperature + Population + Day_temperature:Population + AM + AM:Population, data = GerminationTraitsScaled, family = beta_family(link="logit"))#can use this to check assumptions with DHARMa. The 2 models gives similar output.

#model_performance(Model) ; r.squaredGLMM(Model) #rsq(Model)
Anova(Model)
summary(Model)

Model = betareg(p.beta(GerminationTraitsScaled$GP) ~ Population, data = GerminationTraitsScaled)#clean other effects to keep only population for pairwise comparisons
PairComp_GP = back.emmeans(emmeans(Model, ~ Population, mode="link"),transform="p.beta",n=120) ; colnames(PairComp_GP) = c("Population","SE_inf","EMMean","SE_sup")
PairComp_GP = PairComp_GP %>% arrange(EMMean) %>% left_join(cld(emmeans(Model, ~ Population, mode="link"), adjust = "tukey", Letters = LETTERS)[,c("Population",".group")], by = c("Population")) %>% mutate(Population = factor(Population, levels = unique(Population))) %>% left_join(ClimateProv[,c("Population","ClimaticMoistureDeficit","RM_ThermicityIndex","Tseasonality")],by="Population")

#Cross validation
Germination_training = GerminationTraitsScaled[sample(nrow(GerminationTraitsScaled), 80), ]#training data set
Germination_validate = filter(GerminationTraitsScaled, Id %notin% Germination_training$Id)
ModelCV = betareg(p.beta(Germination_training$GP) ~ Day_temperature + Population + Day_temperature:Population + AM + AM:Population, data = Germination_training)#training of the Model on the training dataset
Predictions = predict(ModelCV, Germination_validate, re.form=NA, type="response")
cor.test(Predictions, Germination_validate$GP)
scatter_plot(obs = Germination_validate$GP, pred = Predictions) + coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))#print obs VS pred on a graphe 1:1

#Output tables
GRGanova_table = rbind(c("GP","","",""),as.data.frame(Anova(Model)) %>% rownames_to_column(., "Covariate") %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "Pr(>Chisq)") %>%  
                         mutate_at(vars(c("Chisq")), function(x) round(as.numeric(x), 2)))
GRGsummary_table = rbind(c("GP","","","",""),as.data.frame(summary(Model)[["coefficients"]])[,c(1:4)] %>% rownames_to_column(., "Covariate") %>%  mutate_all(funs(str_replace(., "Population", ""))) %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "mean.Pr...z..") %>% 
                           mutate_at(vars(c("mean.Estimate","mean.Std..Error","mean.z.value")), function(x) round(as.numeric(x), 2)) %>% rename("Test value" = "mean.z.value","Estimate" = "mean.Estimate","Std. Error" = "mean.Std..Error"))
GRGPerf_table = model_performance(Model) %>% mutate(PearsonsCorrelation_obsVSpred = cor.test(Predictions, Germination_validate$GP)[["estimate"]][["cor"]])
```

```{r  GP specific Population model}
GerminationTraitsScaledGrouped = GerminationTraitsScaled %>% group_by(Population) %>% summarise_all(mean)#no significant problem detected on grouped model

#Model = glmmTMB(p.beta(GerminationTraitsScaled$GP) ~ ClimaticMoistureDeficit + Tmax_spring + Tseasonality + RM_ThermicityIndex, data = GerminationTraitsScaled, family = beta_family(link="logit"))#can also use betareg
#Model = betareg(p.beta(GerminationTraitsScaled$GP) ~ ClimaticMoistureDeficit + Tmax_spring + Tseasonality + RM_ThermicityIndex, data = GerminationTraitsScaled)

#Stepwise selection
#Model = glmmTMB(p.beta(GR) ~  Tseasonality + RM_ThermicityIndex, data = GerminationTraitsScaled, family = beta_family(link="logit"))
Model = betareg(p.beta(GR) ~  Tseasonality + RM_ThermicityIndex, data = GerminationTraitsScaled)

#model_performance(Model) ; r.squaredGLMM(Model) #rsq(Model)
Anova(Model)
summary(Model)

#Cross validation
Germination_training = GerminationTraitsScaled[sample(nrow(GerminationTraitsScaled), 80), ]#training data set
Germination_validate = filter(GerminationTraitsScaled, Id %notin% Germination_training$Id)
ModelCV = betareg(p.beta(GR) ~  Tseasonality + RM_ThermicityIndex, data = Germination_training)#training of the Model on the training dataset
Predictions = predict(ModelCV, Germination_validate, re.form=NA, type="response")
cor.test(Predictions, Germination_validate$GP)
scatter_plot(obs = Germination_validate$GP, pred = Predictions) + coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))#print obs VS pred on a graphe 1:1

#Output tables
GRSanova_table = rbind(c("GP","","",""),as.data.frame(Anova(Model)) %>% rownames_to_column(., "Covariate") %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "Pr(>Chisq)") %>%  
                         mutate_at(vars(c("Chisq")), function(x) round(as.numeric(x), 2)))
GRSsummary_table = rbind(c("GP","","","",""),as.data.frame(summary(Model)[["coefficients"]]$mean) %>% rownames_to_column(., "Covariate") %>%  mutate_all(funs(str_replace(., "Population", ""))) %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "Pr(>|z|)") %>% 
                           mutate_at(vars(c("Estimate","Std. Error","z value")), function(x) round(as.numeric(x), 2)) %>% rename("Test value" = "z value"))
GRSPerf_table = model_performance(Model) %>% mutate(PearsonsCorrelation_obsVSpred = cor.test(Predictions, Germination_validate$GP)[["estimate"]][["cor"]])
```

##GT / AT0 

```{r  Raw data}
ggplot(data = GermIndexes)+
  geom_point(aes(x=Day_temperature,y=AT0 ))+
  geom_smooth(aes(x=Day_temperature,y=AT0 , color = Population))+
  scale_color_manual(values = OakPalette,breaks=TmaxSpring_tab[,c("Population")])+
  theme_bw()

ggplot(data = Germination)+
  geom_point(aes(x=Day_temperature,y=Germination_time))+
  geom_smooth(aes(x=Day_temperature,y=Germination_time, color = Population))+
  scale_color_manual(values = OakPalette,breaks=TmaxSpring_tab[,c("Population")])+
  theme_bw()

ggplot(data = Germination)+
  geom_point(aes(x=Tmax_spring,y=Germination_time))+
  geom_smooth(aes(x=Tmax_spring,y=Germination_time),method = "glm")+
  theme_bw()
```

```{r  Data transformation_remove ungerminated acorns}
GerminationScaledT = filter(GerminationScaled, Germination != 0) %>% #Remove ungerminated accorns
  mutate(Germination_time = ifelse(Germination_time == 0, 1,Germination_time))#germination à 0 n'existe pas ? Correction germination à 1 jour
GerminationTraitsScaledT = filter(GerminationTraitsScaled, !is.na(AT0 ))

GerminationT = filter(Germination, Germination != 0) %>% #Remove ungerminated accorns
  mutate(Germination_time = ifelse(Germination_time == 0, 1,Germination_time))#germination à 0 n'existe pas ? Correction germination à 1 jour
GerminationTraitsT = filter(GerminationTraits, !is.na(AT0 ))
```

Response variable: Germination time -> zero inflated count data
-> Poisson regression (glmer) ? No, because overdispersion of data
-> Negative Binomial regression (glmer.nb) ? Yes, but still no so good
-> Zero inflated Negative Binomial regression (glmmTMB) ? Better

```{r  GT general model}
Model = glmmTMB(Germination_time ~ Day_temperature + Population + Day_temperature:Population + proxi_poids + proxi_poids:Population + (1 | Id_mother), data=GerminationT, ziformula = ~1, family = nbinom1(link = "log"))
AIC(Model)#AIC lower with random effect
model_performance(Model) ; r.squaredGLMM(Model)#Variance expliqué (variance marginale des effets fixes et variance conditionnelle fixe + random)
Anova(Model)
summary(Model)


Model = glmmTMB(Germination_time ~ Day_temperature + Population + Day_temperature:Population + proxi_poids + proxi_poids:Population + (1 | Id_mother), data=filter(GerminationT,Population != "Fernana" & Population != "Moliets et Maa"), ziformula = ~1, family = nbinom1(link = "log"))
effect_plot(Model,pred = "Day_temperature",plot = TRUE,interval = TRUE,int.type = "confidence",int.width = .8)+xlab("Chamber temperature") + ylab("Germination time")+ ylim(0,50)+theme(axis.text=element_text(size=20),axis.title=element_text(size=25),legend.title=element_text(size=20),legend.text=element_text(size=15))

InteractionPlot = interact_plot(Model, pred = "Day_temperature", modx = "Population", plot = TRUE, colors = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")],interval = TRUE, int.type = "confidence", int.width = .8) + scale_linetype_manual(values = rep("solid", length(unique(GerminationT$Population))))+xlab("Chamber temperature") + ylab("Germination time")
InteractionPlot + ylim(0,50)+theme(axis.text=element_text(size=20),axis.title=element_text(size=25),legend.title=element_text(size=20),legend.text=element_text(size=15))

interact_plot(Model, pred = "proxi_poids", modx = "Population", plot = TRUE, colors = OakPalette,interval = TRUE, int.type = "confidence", int.width = .8) + scale_linetype_manual(values = rep("solid", length(unique(GerminationT$Population))))+ scale_color_manual(name = "Population",values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])+xlab("Acorn mass") + ylab("Germination time")

GerminationT[,c("Day_temperature","Germination_time")] %>% group_by(Day_temperature) %>% summarise_all(mean)

#Cross validation
Germination_training = GerminationT[sample(nrow(GerminationT), 400), ]#training data set
Germination_validate = filter(GerminationT, Id_seed %notin% Germination_training$Id_seed) %>% group_by(Population, Day_temperature) %>% mutate(Germination_rate = sum(Germination)/n())
ModelCV = glmmTMB(Germination_time ~ Day_temperature + Population + Day_temperature:Population + proxi_poids + (1 | Id_mother), data=Germination_training, ziformula = ~1, family = nbinom1(link = "log"))#training of the Model on the training dataset
Predictions = predict(ModelCV, Germination_validate, re.form=NA, type="response")
cor.test(Predictions, Germination_validate$Germination_time)
scatter_plot(obs = Germination_validate$Germination_time, pred = Predictions) + coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))#print obs VS pred on a graphe 1:1

#Output tables
GTGanova_table = rbind(c("GT","","",""),as.data.frame(Anova(Model)) %>% rownames_to_column(., "Covariate") %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "Pr(>Chisq)") %>%  
                         mutate_at(vars(c("Chisq")), function(x) round(as.numeric(x), 2)))
GTGsummary_table = rbind(c("GT","","","",""),as.data.frame(summary(Model)[["coefficients"]]$cond) %>% rownames_to_column(., "Covariate") %>%  mutate_all(funs(str_replace(., "Population", ""))) %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "Pr(>|z|)") %>% 
                           mutate_at(vars(c("Estimate","Std. Error","z value")), function(x) round(as.numeric(x), 2)) %>% rename("Test value" = "z value"))
GTGPerf_table = model_performance(Model) %>% mutate(PearsonsCorrelation_obsVSpred = cor.test(Predictions, Germination_validate$Germination_time)[["estimate"]][["cor"]],Model = "General GT model")
```

```{r  GT specific Population model}
#Model = glmmTMB(Germination_time ~ ClimaticMoistureDeficit + Tmax_spring + Tseasonality + RM_ThermicityIndex + Day_temperature + ClimaticMoistureDeficit:Day_temperature + Tmax_spring:Day_temperature + RM_ThermicityIndex:Day_temperature + I(Tseasonality^2):Day_temperature, data=GerminationT, ziformula = ~1, family = nbinom1(link = "log"))
#Stepwise selection
Model = glmmTMB(Germination_time ~ ClimaticMoistureDeficit + Day_temperature + RM_ThermicityIndex + Tmax_spring + Tseasonality + Day_temperature:Tseasonality + (1 | Id_mother), data=GerminationT, ziformula = ~1, family = nbinom1(link = "log"))


model_performance(Model) ; r.squaredGLMM(Model)#Variance expliqué (variance marginale des effets fixes et variance conditionnelle fixe + random)
Anova(Model)
summary(Model)

InteractionPlot = interact_plot(Model, pred = "ClimaticMoistureDeficit", modx = "RM_ThermicityIndex", plot = TRUE, interval = TRUE, int.type = "confidence", int.width = .8) + scale_linetype_manual(values = rep("solid", length(unique(GerminationT$Population))))+ ylab("Germination time")
InteractionPlot + ylim(0,50)+theme(axis.text=element_text(size=20),axis.title=element_text(size=25),legend.title=element_text(size=20),legend.text=element_text(size=15))

#Cross validation
Germination_training = GerminationT[sample(nrow(GerminationT), 400), ]#training data set
Germination_validate = filter(GerminationT, Id_seed %notin% Germination_training$Id_seed) %>% group_by(Population, Day_temperature) %>% mutate(Germination_rate = sum(Germination)/n())
ModelCV = glmmTMB(Germination_time ~ ClimaticMoistureDeficit + Day_temperature + RM_ThermicityIndex + Tmax_spring + Tseasonality + Day_temperature:Tseasonality, data=Germination_training, ziformula = ~1, family = nbinom1(link = "log"))#training of the Model on the training dataset
Predictions = predict(ModelCV, Germination_validate, re.form=NA, type="response")
cor.test(Predictions, Germination_validate$Germination_time)
scatter_plot(obs = Germination_validate$Germination_time, pred = Predictions)#print obs VS pred on a graphe 1:1

#Output tables
GTSanova_table = rbind(c("GT","","",""),as.data.frame(Anova(Model)) %>% rownames_to_column(., "Covariate") %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "Pr(>Chisq)") %>%  
                         mutate_at(vars(c("Chisq")), function(x) round(as.numeric(x), 2)))
GTSsummary_table = rbind(c("GT","","","",""),as.data.frame(summary(Model)[["coefficients"]]$cond) %>% rownames_to_column(., "Covariate") %>%  mutate_all(funs(str_replace(., "Population", ""))) %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "Pr(>|z|)") %>% 
                           mutate_at(vars(c("Estimate","Std. Error","z value")), function(x) round(as.numeric(x), 2)) %>% rename("Test value" = "z value"))
GTSPerf_table = model_performance(Model) %>% mutate(PearsonsCorrelation_obsVSpred = cor.test(Predictions, Germination_validate$Germination_time)[["estimate"]][["cor"]],Model = "Specific GT model")
```

```{r  AT0  general model}
#Model = glm(AT0  ~ Day_temperature + Population + Day_temperature:Population + AM + AM:Population, data = GerminationTraitsScaled, family = poisson(link="identity"))
#Model$deviance / Model$df.residual #>1.5 --> overdispersion --> quasipoisson or negative binomial

Model = glm.nb(AT0  ~ Day_temperature + Population + Day_temperature:Population + AM + AM:Population, data = GerminationTraitsScaledT, link = "identity")
Model$deviance / Model$df.residual #ok

model_performance(Model) ; rsq(Model)
Anova(Model)
summary(Model)

Model = glm.nb(AT0  ~ Day_temperature + Population + Day_temperature:Population, data = GerminationTraitsScaledT, link = "identity")#keep only significant effects for pairwise comparisons
PairComp_AT0  = as.data.frame(emmeans(Model, ~ Population, mode="link")) %>% mutate(SE_inf = emmean - SE,SE_sup = emmean + SE) %>% select(Population,SE_inf,emmean,SE_sup) ; colnames(PairComp_AT0 ) = c("Population","SE_inf","EMMean","SE_sup")
PairComp_AT0  = PairComp_AT0  %>% arrange(EMMean) %>% left_join(cld(emmeans(Model, ~ Population, mode="link"), adjust = "tukey", Letters = LETTERS)[,c("Population",".group")], by = c("Population")) %>% mutate(Population = factor(Population, levels = unique(Population)))%>% left_join(ClimateProv[,c("Population","ClimaticMoistureDeficit","RM_ThermicityIndex","Tseasonality")],by="Population")

Model = glm.nb(AT0  ~ as.factor(Day_temperature):Population, data = GerminationTraitsScaledT, link = "identity")#keep only significant effects for pairwise comparisons
PairComp2_AT0  = as.data.frame(emmeans(Model, ~ Day_temperature | Population, mode="link")) %>% mutate(SE_inf = emmean - SE,SE_sup = emmean + SE,Day_temperature = as.character(ifelse(Day_temperature < 0,"15",ifelse(Day_temperature > 0, "25","20")))) %>% select(Population,Day_temperature,SE_inf,emmean,SE_sup) ; colnames(PairComp2_AT0 ) = c("Population","Day_temperature","SE_inf","EMMean","SE_sup")  
PairComp2_AT0  = PairComp2_AT0  %>% arrange(EMMean) %>% left_join(cld(emmeans(Model, ~ Day_temperature | Population, mode="link"), adjust = "tukey", Letters = LETTERS)[,c("Population","Day_temperature",".group")] %>% mutate(Day_temperature = as.character(ifelse(Day_temperature < 0,"15",ifelse(Day_temperature > 0, "25","20")))), by = c("Population","Day_temperature")) %>% group_by(Population) %>% mutate(Day_temperature = factor(Day_temperature, levels = unique(Day_temperature))) %>% left_join(ClimateProv[,c("Population","ClimaticMoistureDeficit","RM_ThermicityIndex","Tseasonality")],by="Population")

Model = glm.nb(AT0  ~ Day_temperature + Population + Day_temperature:Population + AM + AM:Population, data = GerminationTraitsT, link = "identity")
interact_plot(Model, pred = "Day_temperature", modx = "Population", plot = TRUE, colors = OakPalette,interval = TRUE, int.type = "confidence", int.width = .8) + scale_linetype_manual(values = rep("solid", length(unique(GerminationT$Population))))+ scale_color_manual(name = "Population",values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])+xlab("Chamber temperature") + ylab("Mean germination time")+theme_bw()+ theme(axis.text=element_text(size=30),axis.title=element_text(size=35),
                                                                                                                                                                                                                                                                                                                                                                                                                                                   legend.title = element_text(size = 23),legend.text = element_text(size = 20)) + guides(color = guide_legend(override.aes = list(size = 10)))

Model = glm.nb(AT0  ~ as.factor(Day_temperature) + Population + as.factor(Day_temperature):Population + AM + AM:Population, data = GerminationTraitsT %>% mutate(Day_temperature=as.factor(Day_temperature)), link = "identity")
interact_plot(Model, pred = "AM", modx = "Day_temperature", plot = TRUE, colors = c("15" = "#3399FF", "20" = "#009900", "25" = "#CC3333"),interval = TRUE, int.type = "confidence", int.width = .8) + scale_linetype_manual(values = rep("solid", length(unique(GerminationT$Population))))+theme_bw()+theme(axis.text=element_text(size=30),axis.title=element_text(size=35),
                                                                                                                                                                                                                                                                                                             legend.title = element_text(size = 23),legend.text = element_text(size = 20)) + guides(color = guide_legend(override.aes = list(size = 10)))

#Cross validation
Germination_training = GerminationTraitsScaledT[sample(nrow(GerminationTraitsScaledT), 80), ]#training data set
Germination_validate = filter(GerminationTraitsScaledT, Id %notin% Germination_training$Id)
ModelCV = glm.nb(AT0  ~ Day_temperature + Population + Day_temperature:Population + AM + AM:Population, data = Germination_training, link = "identity")#training of the Model on the training dataset
Predictions = predict(ModelCV, Germination_validate, re.form=NA, type="response")
cor.test(Predictions, Germination_validate$AT0 )
scatter_plot(obs = Germination_validate$AT0 , pred = Predictions)#print obs VS pred on a graphe 1:1

#Output tables
AT0 Ganova_table = rbind(c("AT0","","",""),as.data.frame(Anova(Model)) %>% rownames_to_column(., "Covariate") %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "Pr(>Chisq)") %>%  
                          mutate_at(vars(c("LR Chisq")), function(x) round(as.numeric(x), 2))) %>% rename(Chisq = "LR Chisq")
AT0 Gsummary_table = rbind(c("AT0","","","",""),as.data.frame(summary(Model)[["coefficients"]]) %>% rownames_to_column(., "Covariate") %>%  mutate_all(funs(str_replace(., "Population", ""))) %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "Pr(>|z|)") %>% 
                            mutate_at(vars(c("Estimate","Std. Error","z value")), function(x) round(as.numeric(x), 2)) %>% rename("Test value" = "z value"))
AT0 GPerf_table = model_performance(Model) %>% mutate(PearsonsCorrelation_obsVSpred = cor.test(Predictions, Germination_validate$AT0 )[["estimate"]][["cor"]])
```

```{r  AT0  specific Population model}
Model = glm.nb(AT0  ~ ClimaticMoistureDeficit + Tmax_spring + Tseasonality + RM_ThermicityIndex + Day_temperature +  ClimaticMoistureDeficit:Day_temperature + Tmax_spring:Day_temperature + RM_ThermicityIndex:Day_temperature + Tseasonality:Day_temperature, data = GerminationTraitsScaledT, link = "identity")
Model$deviance / Model$df.residual #ok
#Stepwise selection: remove interaction, doesn't change the result

model_performance(Model) ; rsq(Model)
Anova(Model)
summary(Model)

#Cross validation
Germination_training = GerminationTraitsScaledT[sample(nrow(GerminationTraitsScaledT), 80), ]#training data set
Germination_validate = filter(GerminationTraitsScaledT, Id %notin% Germination_training$Id)
ModelCV = glm.nb(AT0  ~ ClimaticMoistureDeficit + Tmax_spring + Tseasonality + RM_ThermicityIndex + Day_temperature + ClimaticMoistureDeficit:Day_temperature + Tmax_spring:Day_temperature + RM_ThermicityIndex:Day_temperature + Tseasonality:Day_temperature, data = Germination_training, link = "identity")#training of the Model on the training dataset
Predictions = predict(ModelCV, Germination_validate, re.form=NA, type="response")
cor.test(Predictions, Germination_validate$AT0 )
scatter_plot(obs = Germination_validate$AT0 , pred = Predictions,na.rm = T)#print obs VS pred on a graphe 1:1

#Output tables
AT0 Sanova_table = rbind(c("AT0","","",""),as.data.frame(Anova(Model)) %>% rownames_to_column(., "Covariate") %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "Pr(>Chisq)") %>%  
                          mutate_at(vars(c("LR Chisq")), function(x) round(as.numeric(x), 2))) %>% rename(Chisq = "LR Chisq")
AT0 Ssummary_table = rbind(c("AT0","","","",""),as.data.frame(summary(Model)[["coefficients"]]) %>% rownames_to_column(., "Covariate") %>%  mutate_all(funs(str_replace(., "Population", ""))) %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "Pr(>|z|)") %>% 
                            mutate_at(vars(c("Estimate","Std. Error","z value")), function(x) round(as.numeric(x), 2)) %>% rename("Test value" = "z value"))
AT0 SPerf_table = model_performance(Model) %>% mutate(PearsonsCorrelation_obsVSpred = cor.test(Predictions, Germination_validate$AT0 )[["estimate"]][["cor"]])
```

##GS

```{r  GS general model}
#GS close to 1 = very synchronized
ggplot(data = GerminationTraits)+geom_histogram(aes(GS))#quantitative continuous between 0 an 1 (proportion): beta regression
ggplot(data = GerminationTraits)+geom_histogram(aes(p.beta(GerminationTraits$GS)))#put it in ]0; 1[ instead of [0; 1] 

Model = betareg(p.beta(GerminationTraitsScaledT$GS) ~ Day_temperature + Population + Day_temperature:Population + AM + AM:Population, data = GerminationTraitsScaledT)
#Model = glmmTMB(p.beta(GerminationTraitsScaledT$GS) ~ Day_temperature + Population + Day_temperature:Population + AM + AM:Population, data = GerminationTraitsScaledT, family = beta_family(link="logit"))#can use this to check assumptions with DHARMa. The 2 models gives similar output.

#model_performance(Model) ; r.squaredGLMM(Model) #rsq(Model)
Anova(Model)
summary(Model)

Model = betareg(p.beta(GerminationTraitsScaledT$GS) ~  Population, data = GerminationTraitsScaledT)#clean other effect to keep only population for pairwise comparisons
PairComp_GS = back.emmeans(emmeans(Model, ~ Population, mode="link"),transform="p.beta",n=120) ; colnames(PairComp_GS) = c("Population","SE_inf","EMMean","SE_sup")
PairComp_GS = PairComp_GS %>% arrange(EMMean) %>% left_join(cld(emmeans(Model, ~ Population, mode="link"), adjust = "tukey", Letters = LETTERS)[,c("Population",".group")], by = c("Population")) %>% mutate(Population = factor(Population, levels = unique(Population)))%>% left_join(ClimateProv[,c("Population","ClimaticMoistureDeficit","RM_ThermicityIndex","Tseasonality")],by="Population")

#Cross validation
Germination_training = GerminationTraitsScaledT[sample(nrow(GerminationTraitsScaledT), 80), ]#training data set
Germination_validate = filter(GerminationTraitsScaledT, Id %notin% Germination_training$Id)
ModelCV = betareg(p.beta(Germination_training$GS) ~ Day_temperature + Population + Day_temperature:Population + AM + AM:Population, data = Germination_training)#training of the Model on the training dataset
Predictions = predict(ModelCV, Germination_validate, re.form=NA, type="response")
cor.test(Predictions, Germination_validate$GS)
scatter_plot(obs = Germination_validate$GS, pred = Predictions) + coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))#print obs VS pred on a graphe 1:1

#Output tables
GSGanova_table = rbind(c("GS","","",""),as.data.frame(Anova(Model)) %>% rownames_to_column(., "Covariate") %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "Pr(>Chisq)") %>%  
                         mutate_at(vars(c("Chisq")), function(x) round(as.numeric(x), 2)))
GSGsummary_table = rbind(c("GS","","","",""),as.data.frame(summary(Model)[["coefficients"]])[1:4] %>% rownames_to_column(., "Covariate") %>%  mutate_all(funs(str_replace(., "Population", ""))) %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "mean.Pr...z..") %>% 
                           mutate_at(vars(c("mean.Estimate","mean.Std..Error","mean.z.value")), function(x) round(as.numeric(x), 2)) %>% rename("Test value" = "mean.z.value","Estimate" = "mean.Estimate","Std. Error" = "mean.Std..Error"))
GSGPerf_table = model_performance(Model) %>% mutate(PearsonsCorrelation_obsVSpred = cor.test(Predictions, Germination_validate$GS)[["estimate"]][["cor"]],Model = "General GS model")
```

```{r  GS specific Population model}
#Model = glmmTMB(p.beta(GerminationTraitsScaledT$GS) ~ ClimaticMoistureDeficit + Tmax_spring + Tseasonality + RM_ThermicityIndex, data = GerminationTraitsScaledT, family = beta_family(link="logit"))

#Stepwise selection
#Model = glmmTMB(p.beta(GerminationTraitsScaledT$GS) ~ RM_ThermicityIndex + Tseasonality, data = GerminationTraitsScaledT, family = beta_family(link="logit"))
Model = betareg(p.beta(GerminationTraitsScaledT$GS) ~ RM_ThermicityIndex + Tseasonality, data = GerminationTraitsScaledT)
Model = glmmTMB(p.beta(GerminationTraitsScaledT$GS) ~ RM_ThermicityIndex + Tseasonality, data = GerminationTraitsScaledT, family = beta_family())

model_performance(Model) ; r.squaredGLMM(Model) #rsq(Model)
Anova(Model)
summary(Model)

#Cross validation
Germination_training = GerminationTraitsScaledT[sample(nrow(GerminationTraitsScaledT), 80), ]#training data set
Germination_validate = filter(GerminationTraitsScaledT, Id %notin% Germination_training$Id)
Model = betareg(p.beta(Germination_training$GS) ~ RM_ThermicityIndex + Tseasonality, data = Germination_training)#training of the Model on the training dataset
Predictions = predict(ModelCV, Germination_validate, re.form=NA, type="response")
cor.test(Predictions, Germination_validate$GS)
scatter_plot(obs = Germination_validate$GS, pred = Predictions)#print obs VS pred on a graphe 1:1

#Output tables
GSSanova_table = rbind(c("GS","","",""),as.data.frame(Anova(Model)) %>% rownames_to_column(., "Covariate") %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "Pr(>Chisq)") %>%  
                         mutate_at(vars(c("Chisq")), function(x) round(as.numeric(x), 2)))
GSSsummary_table = rbind(c("GS","","","",""),as.data.frame(summary(Model)[["coefficients"]])[1:4] %>% rownames_to_column(., "Covariate") %>%  mutate_all(funs(str_replace(., "Population", ""))) %>% mutate_all(funs(str_replace(., "Day_temperature", "Chamber temperature"))) %>% mutate_all(funs(str_replace(., "proxi_poids", "Acorn Mass"))) %>% rename(p = "mean.Pr...z..") %>% 
                           mutate_at(vars(c("mean.Estimate","mean.Std..Error","mean.z.value")), function(x) round(as.numeric(x), 2)) %>% rename("Test value" = "mean.z.value","Estimate" = "mean.Estimate","Std. Error" = "mean.Std..Error"))
GSSPerf_table = model_performance(Model) %>% mutate(PearsonsCorrelation_obsVSpred = cor.test(Predictions, Germination_validate$GS)[["estimate"]][["cor"]],Model = "Specific GS model")
```

##Tables export

```{r  Tables export}
#Anova Table
##General model Table
Ganova_table = condformat(do.call(rbind,args = list(GGanova_table, GRGanova_table, GTGanova_table, AT0 Ganova_table, GSGanova_table, CvGTGanova_table,CoxGanova_table))) %>% rule_text_bold(c("Covariate","Chisq","Df","p"), p <= 0.05, na.bold = FALSE, lockcells = FALSE) %>% mutate(p = ifelse(as.numeric(p) < 0.001,"<0.001",round(as.numeric(p),3))) 
##Specific model table
Sanova_table = condformat(do.call(rbind,args = list(GSanova_table, GRSanova_table, GTSanova_table, AT0 Sanova_table, GSSanova_table, CvGTSanova_table,CoxSanova_table))) %>% rule_text_bold(c("Covariate","Chisq","Df","p"), p <= 0.05, na.bold = FALSE, lockcells = FALSE) %>% mutate(p = ifelse(as.numeric(p) < 0.001,"<0.001",round(as.numeric(p),3)))

#Summary Table
##General model Table
Gsummary_table = condformat(do.call(rbind,args = list(GGsummary_table, GRGsummary_table, GTGsummary_table, AT0 Gsummary_table, GSGsummary_table, CvGTGsummary_table,CoxGsummary_table))) %>% rule_text_bold(c("Covariate","Estimate","Std. Error","Test value","p"), p <= 0.05, na.bold = FALSE, lockcells = FALSE) %>% mutate(p = ifelse(as.numeric(p) < 0.001,"<0.001",round(as.numeric(p),3)))
##Specific model table
Ssummary_table = condformat(do.call(rbind,args = list(GSsummary_table, GRSsummary_table, GTSsummary_table, AT0 Ssummary_table, GSSsummary_table, CvGTSsummary_table,CoxSsummary_table))) %>% rule_text_bold(c("Covariate","Estimate","Std. Error","Test value","p"), p <= 0.05, na.bold = FALSE, lockcells = FALSE) %>% mutate(p = ifelse(as.numeric(p) < 0.001,"<0.001",round(as.numeric(p),3)))

#Performance Table
##General model Table
GPerf_table = condformat(do.call(plyr::rbind.fill,args = list(GGPerf_table, GTGPerf_table, GSGPerf_table, CoxGPerf_table)) %>% mutate(across(where(is.numeric), round, 3)))#GRGPerf_table AT0 GPerf_table
##Specific model table
SPerf_table = condformat(do.call(plyr::rbind.fill,args = list(GSPerf_table, GTSPerf_table, GSSPerf_table,CoxSPerf_table))%>% mutate(across(where(is.numeric), round, 3)))#GRSPerf_table AT0 SPerf_table

#Cox PH table
PH_table = condformat(do.call(plyr::rbind.fill,args = list(GPH_table,SPH_table)) %>% mutate(across(where(is.numeric), round, 3)))

#Export
condformat2excel(Ganova_table, file="ModelsTableOutputs.xlsx", sheet_name="Ganova_table", overwrite_wb = FALSE, overwrite_sheet = TRUE)
condformat2excel(Sanova_table, file="ModelsTableOutputs.xlsx", sheet_name="Sanova_table", overwrite_wb = FALSE, overwrite_sheet = TRUE)
condformat2excel(Gsummary_table, file="ModelsTableOutputs.xlsx", sheet_name="Gsummary_table", overwrite_wb = FALSE, overwrite_sheet = TRUE)
condformat2excel(Ssummary_table, file="ModelsTableOutputs.xlsx", sheet_name="Ssummary_table", overwrite_wb = FALSE, overwrite_sheet = TRUE)
condformat2excel(GPerf_table, file="ModelsTableOutputs.xlsx", sheet_name="GPerf_table", overwrite_wb = FALSE, overwrite_sheet = TRUE)
condformat2excel(SPerf_table, file="ModelsTableOutputs.xlsx", sheet_name="SPerf_table", overwrite_wb = FALSE, overwrite_sheet = TRUE)
condformat2excel(times_table, file="ModelsTableOutputs.xlsx", sheet_name="CoxTimes", overwrite_wb = FALSE, overwrite_sheet = TRUE)
condformat2excel(PH_table, file="ModelsTableOutputs.xlsx", sheet_name="CoxPH", overwrite_wb = FALSE, overwrite_sheet = TRUE)
```

#Germination indexes models graphical representation 

Summary of the genetic effects (i.e. population climatic variables), environment effects (i.e. chamber temperature), and their interaction (i.e. chamber temperature effect according to population) on germination metrics

```{r  Boxplot and curves Traits according to population}
library(boot)# Pour inv.logit()
GerminationTraits_stats = GerminationTraits %>% 
  group_by(Population) %>% 
  summarise_at(vars(GP,AT0 ,GS),
               list(mean = mean,
                    sd = sd),na.rm=T)

# Ensure Population is a factor with levels in thermicity order
PairComp_GP$Population <- factor(PairComp_GP$Population, levels = RM_ThermicityIndex_tab$Population)
GerminationTraits$Population <- factor(GerminationTraits$Population, levels = RM_ThermicityIndex_tab$Population)
ggplot()+
  geom_point(aes(x=Population, y=EMMean, color=Population, fill=Population), data=PairComp_GP, shape=21, size=4) +
  geom_jitter(aes(x=Population, y=GP, color=Population, alpha=0.2), data=GerminationTraits) +
  geom_errorbar(aes(x=Population, ymin=SE_inf, ymax=SE_sup, color=Population), data=PairComp_GP, size=2) +
  geom_text(aes(x=Population, label=PairComp_GP$.group, color=Population, y=0.25), show.legend=FALSE, size=10, data=PairComp_GP) +
  ylab("Germination \n percentage") +
  theme_bw() +
  theme(axis.text.y=element_text(size=30), axis.title.y=element_text(size=35), axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), legend.title=element_text(size=18), legend.text=element_text(size=15)) +
  scale_color_manual(values=OakPalette, breaks=RM_ThermicityIndex_tab$Population) +
  scale_fill_manual(values=OakPalette, breaks=RM_ThermicityIndex_tab$Population)

ggplot()+
  geom_errorbar(aes(x=RM_ThermicityIndex,ymin = SE_inf, ymax = SE_sup, color=Population), data=PairComp_GP, size=2)+
  geom_point(aes(x=RM_ThermicityIndex,y=EMMean, fill=Population),data=PairComp_GP,shape = 21,size = 4)+
  geom_smooth(aes(x=RM_ThermicityIndex,y=EMMean),data=PairComp_GP,method="auto", se=TRUE, fullrange=T,span = 1.5,color = "black", fill = "#FFCC99")+
  theme_bw()+ylab("Germination rate")+xlab("Rivas-Martínez Thermicity Index")+
  theme(axis.text=element_text(size=30),axis.title=element_text(size=35))+
  scale_color_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])+
  scale_fill_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])
ggplot()+
  geom_errorbar(aes(x=Tseasonality,ymin = SE_inf, ymax = SE_sup, color=Population), data=PairComp_GP, size=2)+
  geom_point(aes(x=Tseasonality,y=EMMean, fill=Population),data=PairComp_GP,shape = 21,size = 4)+
  geom_smooth(aes(x=Tseasonality,y=EMMean),data=PairComp_GP,method="auto", se=TRUE, fullrange=T,span = 1.5,color = "black", fill = "#FFCC99")+
  theme_bw()+ylab("Germination rate")+xlab("Temperature Seasonality")+
  theme(axis.text=element_text(size=30),axis.title=element_text(size=35))+
  scale_color_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])+
  scale_fill_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])

# Ensure Population is a factor with levels in thermicity order
PairComp_AT0$Population <- factor(PairComp_AT0$Population, levels = RM_ThermicityIndex_tab$Population)
GerminationTraits$Population <- factor(GerminationTraits$Population, levels = RM_ThermicityIndex_tab$Population)
ggplot()+
  geom_point(aes(x=Population,y=EMMean,color=Population,fill = Population),data=PairComp_AT0 ,shape = 21,size = 4)+
  geom_jitter(aes(x=Population,y=AT0 ,color=Population,alpha=0.2),data=GerminationTraits)+
  geom_errorbar(aes(x=Population,ymin = SE_inf, ymax = SE_sup, color=Population), data=PairComp_AT0 , size=2)+
  geom_text(aes(x = Population,label = PairComp_AT0 $.group, color=Population,y=40), show.legend=FALSE, size = 10,data = PairComp_AT0 )+ ylab("Averaged \n germination time")+
  theme_bw()+
  theme(axis.text.y=element_text(size=30),axis.title.y=element_text(size=35),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank(),legend.title=element_text(size=18),legend.text=element_text(size=15))+
  scale_color_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])+
  scale_fill_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])

ggplot()+
  geom_point(aes(x=RM_ThermicityIndex,y=EMMean, fill=Population),data=PairComp_AT0 ,shape = 21,size = 4)+
  geom_smooth(aes(x=RM_ThermicityIndex,y=EMMean),data=PairComp_AT0 ,method="auto", se=TRUE, fullrange=T,span = 1.5,color = "black", fill = "#FFCC99")+
  geom_errorbar(aes(x=RM_ThermicityIndex,ymin = SE_inf, ymax = SE_sup, color=Population), data=PairComp_AT0 , size=2)+
  theme_bw()+ylab("Mean germination time")+xlab("Rivas-Martínez Thermicity Index")+
  theme(axis.text=element_text(size=30),axis.title=element_text(size=35))+
  scale_color_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])+
  scale_fill_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])
ggplot()+
  geom_point(aes(x=ClimaticMoistureDeficit,y=EMMean, fill=Population),data=PairComp_AT0 ,shape = 21,size = 4)+
  geom_smooth(aes(x=ClimaticMoistureDeficit,y=EMMean),data=PairComp_AT0 ,method="auto", se=TRUE, fullrange=T,span = 1.5,color = "black", fill = "#FFCC99")+
  geom_errorbar(aes(x=ClimaticMoistureDeficit,ymin = SE_inf, ymax = SE_sup, color=Population), data=PairComp_AT0 , size=2)+
  theme_bw()+ylab("Mean germination time")+xlab("Climatic Moisture Deficit")+
  theme(axis.text=element_text(size=30),axis.title=element_text(size=35))+
  scale_color_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])+
  scale_fill_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])
ggplot()+
  geom_point(aes(x=Tseasonality,y=EMMean, fill=Population),data=PairComp_AT0 ,shape = 21,size = 4)+
  geom_smooth(aes(x=Tseasonality,y=EMMean),data=PairComp_AT0 ,method="auto", se=TRUE, fullrange=T,span = 1.5,color = "black", fill = "#FFCC99")+
  geom_errorbar(aes(x=Tseasonality,ymin = SE_inf, ymax = SE_sup, color=Population), data=PairComp_AT0 , size=2)+
  theme_bw()+ylab("Mean germination time")+xlab("Temperature Seasonality")+
  theme(axis.text=element_text(size=30),axis.title=element_text(size=35))+
  scale_color_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])+
  scale_fill_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])
ggplot()+
  geom_smooth(aes(x=Tseasonality,y=EMMean,color=Day_temperature, fill = Day_temperature),data=PairComp2_AT0 ,method="auto", se=TRUE, fullrange=T,span = 1.5, alpha=0.2)+
  geom_point(aes(x=Tseasonality,y=EMMean,color=Day_temperature),data=PairComp2_AT0 ,size = 2, position = position_dodge(width=20))+
  geom_errorbar(aes(x=Tseasonality,ymin = SE_inf, ymax = SE_sup,color=Day_temperature), data=PairComp2_AT0 , size=1, position = position_dodge(width=20))+
  theme_bw()+ylab("Mean germination time")+xlab("Temperature Seasonality")+
  theme(axis.text=element_text(size=30),axis.title=element_text(size=35),legend.text = element_text(size = 16),legend.title = element_text(size = 18))+
  scale_color_manual(values = c("15" = "#0072B2", "20" = "#F0E442", "25" = "#D55E00"),name = "Chamber temperature")+
  scale_fill_manual(values = c("15" = "#0072B2", "20" = "#F0E442", "25" = "#D55E00"),name = "Chamber temperature")

# Ensure Population is a factor with levels in thermicity order
PairComp_GS$Population <- factor(PairComp_GS$Population, levels = RM_ThermicityIndex_tab$Population)
GerminationTraits$Population <- factor(GerminationTraits$Population, levels = RM_ThermicityIndex_tab$Population)
ggplot()+
  geom_point(aes(x=Population,y=EMMean,color=Population,fill = Population),data=PairComp_GS,shape = 21,size = 4)+
  geom_jitter(aes(x=Population,y=GS,color=Population,alpha=0.2),data=GerminationTraits)+
  geom_errorbar(aes(x=Population,ymin = SE_inf, ymax = SE_sup, color=Population), data=PairComp_GS, size=2)+
  geom_text(aes(x = Population,label = PairComp_GS$.group, color=Population,y=0.8), size = 10,data = PairComp_GS)+ ylab("Germination \n synchrony")+ylim(0,1)+
  theme_bw()+
  theme(axis.text.y=element_text(size=30),axis.title.y=element_text(size=35),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank(),legend.title=element_text(size=18),legend.text=element_text(size=15))+
  scale_color_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])+
  scale_fill_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])


ggplot()+
  geom_point(aes(x=Population,y=EMMean,color=Population,fill = Population),data=PairComp_AT0 ,shape = 21,size = 4)+
  geom_jitter(aes(x=Population,y=AT0 ,color=Population,alpha=0.2),data=GerminationTraits)+
  geom_errorbar(aes(x=Population,ymin = SE_inf, ymax = SE_sup, color=Population), data=PairComp_AT0 , size=2)+
  geom_text(aes(x = Population,label = PairComp_AT0 $.group, color=Population,y=40), show.legend=FALSE, size = 10,data = PairComp_AT0 )+ ylab("Averaged \n germination time")+
  theme_bw()+
  theme(axis.text.y=element_text(size=30),axis.title.y=element_text(size=35),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank(),legend.title=element_text(size=18),legend.text=element_text(size=15))+
  scale_color_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])+
  scale_fill_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])

ggplot()+
  geom_point(aes(x=RM_ThermicityIndex,y=EMMean, fill=Population),data=PairComp_GS,shape = 21, size = 4)+
  geom_smooth(aes(x=RM_ThermicityIndex,y=EMMean),data=PairComp_GS,method="auto", se=TRUE, fullrange=T,span = 1.5,color = "black", fill = "#FFCC99")+
  geom_errorbar(aes(x=RM_ThermicityIndex,ymin = SE_inf, ymax = SE_sup, color=Population), data=PairComp_GS, size=2)+
  theme_bw()+ylab("Germination synchrony")+xlab("Rivas-Martínez Thermicity Index")+
  theme(axis.text=element_text(size=30),axis.title=element_text(size=35),legend.text = element_text(size = 16),legend.title = element_text(size = 18))+
  scale_color_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])+
  scale_fill_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])

ggplot()+
  geom_point(aes(x=RM_ThermicityIndex,y=EMMean, fill=Population),data=PairComp_GS,shape = 21,size = 4)+
  geom_smooth(aes(x=RM_ThermicityIndex,y=EMMean),data=PairComp_GS,method="auto", se=TRUE, fullrange=T,span = 1.5,color = "black", fill = "#FFCC99")+
  geom_errorbar(aes(x=RM_ThermicityIndex,ymin = SE_inf, ymax = SE_sup, color=Population), data=PairComp_GS, size=2)+
  theme_bw()+ylab("Germination synchrony")+xlab("Rivas-Martínez Thermicity Index")+
  theme(axis.text=element_text(size=30),axis.title=element_text(size=35),legend.text = element_text(size = 16),legend.title = element_text(size = 18))+
  scale_color_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])+
  scale_fill_manual(values = OakPalette, breaks=RM_ThermicityIndex_tab[,c("Population")])

GerminationTraits_stats2 = GerminationTraits[,c("Population","Day_temperature","AT0")] %>% 
  mutate(Day_temperature = as.factor(Day_temperature)) %>% 
  group_by(Population,Day_temperature) %>% 
  summarise_at(vars(AT0 ),
               list(mean = mean,
                    sd = sd),na.rm=T) %>% 
  left_join(RM_Tseason_tab,by="Population")

ggplot()+
  geom_point(aes(x=reorder(Population, Tseasonality),y=EMMean, group = reorder_within(Day_temperature, EMMean, Population), color=Day_temperature, fill = Day_temperature), data=PairComp2_AT0 ,shape = 21,size = 4, position = position_dodge(width=0.9))+
  geom_jitter(aes(x=reorder(Population, AT0 ),y=AT0 , group = reorder_within(Day_temperature, AT0 , Population), color=Day_temperature, fill = Day_temperature,alpha=0.2),data=GerminationTraits %>% mutate(Day_temperature=as.character(Day_temperature)), position = position_dodge(width=0.9))+
  geom_errorbar(aes(x=reorder(Population, Tseasonality),ymin = SE_inf, ymax = SE_sup, group = reorder_within(Day_temperature, EMMean, Population), color=Day_temperature), data=PairComp2_AT0 , size=2, position = position_dodge(width=0.9))+
  geom_text(aes(x=reorder(Population, Tseasonality),label = PairComp2_AT0 $.group, group = reorder_within(Day_temperature, EMMean, Population), color=Day_temperature,y=40), size = 15,data = PairComp2_AT0 , position = position_dodge(width=0.9))+ ylab("Mean germination \n time")+
  theme_bw()+
  theme(axis.text.y=element_text(size=30),axis.text.x=element_text(size=30,angle = -10), axis.title.y=element_text(size=35), axis.title.x=element_blank(), axis.ticks.x=element_blank(),legend.title=element_text(size=20),legend.text=element_text(size=15))+
  scale_color_manual(values = c("15" = "#0072B2", "20" = "#F0E442", "25" = "#D55E00"),name = "Chamber temperature")+
  scale_fill_manual(values = c("15" = "#0072B2", "20" = "#F0E442", "25" = "#D55E00"),name = "Chamber temperature")
```

```{r  Patterns of germination in time}
ggplot(data = GerminationT)+
  geom_histogram(aes(x = Germination_time))+
  theme_bw()+
  facet_wrap2( ~ factor(Population, levels=c("Tabarka", "Los Alcornocales", "Fernana", "Quinta de Serra", "Fuencaliente", "Tuscania", "Moliets et Maa", "Padouen")), ncol=4, nrow = 2,strip = strip_themed(background_x = elem_list_rect(fill = OakPalette)))
```


```{r  ACP G(0,1) & GR}
ACP = PCA(BothScaled[,c("Population","GP","Germination","Tseasonality","RM_ThermicityIndex")], scale.unit =TRUE, ncp = 6, graph = F, quali.sup = c(1,3))

fviz_pca_var(ACP, axes = c(1, 2), col.quanti.sup = "blue", col.var = factor(c("Trait", "Temperature seasonality", "Cold intensity and \n annual temperature")),palette = ACPVarPalette,legend.title = list(color = "Type"),repel=T)

fviz_eig(ACP, addlabels = TRUE, ylim = c(0, 50))#Graphe des éboulis
```

```{r  ACP GT & AT0 }
ACP = PCA(filter(BothScaled[,c("Population","AT0","Germination_time","ClimaticMoistureDeficit","RM_ThermicityIndex","Tmax_spring","Tseasonality")],!is.na(Germination_time) & !is.na(AT0 )), scale.unit =TRUE, ncp = 8, graph = F, quali.sup = c(1))

fviz_pca_var(ACP, axes = c(1, 2), col.quanti.sup = "blue", col.var = factor(c("Trait", "Trait","Water stress","Cold intensity and \n annual temperature","Spring temperature","Temperature seasonality")),palette = ACPVarPalette,legend.title = list(color = "Type"),repel=T)
fviz_pca_var(ACP, axes = c(1, 3), col.quanti.sup = "blue", col.var = factor(c("Trait", "Trait","Water stress","Cold intensity and \n annual temperature","Spring temperature","Temperature seasonality")),palette = ACPVarPalette,legend.title = list(color = "Type"),repel=T)

fviz_eig(ACP, addlabels = TRUE, ylim = c(0, 50))#Graphe des éboulis
```

```{r  ACP GS}
ACP = PCA(filter(BothScaled[,c("Population","GS","RM_ThermicityIndex")],!is.na(GS)), scale.unit =TRUE, ncp = 8, graph = F, quali.sup = c(1))

fviz_pca_var(ACP, axes = c(1, 2), col.quanti.sup = "blue", col.var = factor(c("Trait", "Cold intensity and \n annual temperature")),palette = ACPVarPalette,legend.title = list(color = "Type"),repel=T)

fviz_eig(ACP, addlabels = TRUE, ylim = c(0, 50))#Graphe des éboulis
```

Aide-mémoire de l’écriture des effets aléatoires dans un modèle mixte 
Voir https://eshinjolly.com/pymer4/rfx_cheatsheet.html
Random intercepts only
(1 | Group)
Random slopes only
(0 + Variable | Group)
Random intercepts and slopes (and their correlation)
(Variable | Group) (modèle plus général que celui juste en dessous)
Random intercepts and slopes (without their correlation)
(1 | Group) + (0 + Variable | Group)
Same as above but will not separate factors (seK: https://rdrr.io/cran/lme4/man/expandDoubleVerts.html)
(Variable || Group)
Random intercept and slope for more than one variable (and their correlations)
(Variable_1 + Variable_2 | Group)
